
## Executive Summary {-}

### Stock Description {-}

This assessment reports the status of the `r spp` (*Anoplopoma fimbria*, or `black cod') resource off the coast of the United States (U.S.) from southern California to the U.S.-Canadian border using data through `r data_end_year`. The resource is modeled as a single stock, however `r spp` do disperse to and from offshore sea mounts and along the coastal waters of the continental U.S., Canada, and Alaska and across the Aleutian Islands to the western Pacific. Their movement is not explicitly accounted for in this analysis.

### Catches {-}

Check to see if referencing works @tbl-es-catches


```{r, results = "asis"}
#| label: tbl-es-catches
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) catches_es_cap 

catches_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:6,
    decimals = 0
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```


### Data and Assessments {-}


### Stock Output and Dynamics {-}

@tbl-es-sb

```{r, results = "asis"}
#| label: tbl-es-sb
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) ssb_es_cap 

ssb_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12  
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```


### Recruitment {-}

```{r, results = "asis"}
#| label: tbl-es-recr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recr_es_cap 

recr_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```


### Exploitation Status {-}


```{r, results = "asis"}
#| label: tbl-es-spr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) spr_es_cap 

spr_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(80)
  ) |>
  as_latex()

```


### Ecosysystem Consideration {-}


### Reference Points {-}


```{r, results = "asis"}
#| label: tbl-ref-points
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) reference_points_cap 


reference_points_table |>
  gt() |>
  fmt_number(
    columns = 2:4,
    rows = c(1:4, 7, 10, 12, 15, 17, 20),
    decimals = 0
  ) |>
  fmt_number(
    columns = 2:4,
    rows = c(5, 8, 9, 13:14, 18:19),
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>  
  tab_style(
    style = list(
      cell_text(style = "italic")
    ),
    locations = cells_body(
      columns = "Reference Point",
      rows = starts_with("Reference")
    )
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "---"
  ) |>
  cols_width(
    starts_with("R") ~ px(300),
    starts_with("E") ~ px(65),
    ends_with("l") ~ px(65)
  ) |>
  as_latex()

```

### Management Performance {-}


```{r, results = "asis"}
#| label: tbl-es-management
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recent_management_cap 

recent_management_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 5:6,
    decimals = 0
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```


### Evaluation of Scientific Uncertainty {-}


### Harvest Projections and Decision Tables {-}

\blandscape

```{r, results = "asis"}
#| label: tbl-es-projections
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) projections_cap 

projections_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(2:3, 5),
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = 6,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```

\elandscape

### Unresolved Problems and Major Uncertainties {-}

### Research and Data Needs {-}

