
# Executive Summary {-}

## Stock Description {-}

This assessment reports the status of the `r spp` (*Anoplopoma fimbria*, or `black cod') resource off the coast of the United States (U.S.) from southern California to the U.S.-Canadian border using data through `r data_end_year`. The resource is modeled as a single stock, however `r spp` do disperse to and from offshore sea mounts and along the coastal waters of the continental U.S., Mexico, Canada, and Alaska and across the Aleutian Islands to the western Pacific. Net changes in population size or structure due to movement across the assessment boundary are considered negligible, though alternative assumptions are explored using sensitivity model evaluations.

## Catches {-}

Check to see if referencing works @tbl-es-catches


```{r, results = "asis"}
#| label: tbl-es-catches
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) catches_es_cap 
#| tbl-pos: H

catches_es_table |>
  dplyr::rename(
    `TWL Discards (mt)` = `TWL_Discards (mt)`,
    `HKL Discards (mt)` = `HKL_Discards (mt)`,
    `Pot Discards (mt)` = `Pot_Discard (mt)`
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:8,
    decimals = 0
  ) |>
  gt::tab_options(
    table.font.size = 12,
    # if want to not have tables the width of the page include the below option:
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  gt::as_latex()

```

```{r} 
#| label: fig-es-landings
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Landings in metric tons (mt) by year for trawl, pot, and hook-and line gears."
#| fig-alt: "Landings before the 1970s were less than 5,000 mt by year with landings sharply increasing in the 1970s and early 1980s and landings since ranging between approximately 5,000 to 7,500 mt by year."

knitr::include_graphics(
  file.path(model_dir, "plots", "catch2_landings_stacked.png")
)

```

## Data and Assessments {-}


## Stock Output and Dynamics {-}

For the purposes of this assessment, female spawning biomass is assumed to be proportional to egg and larval production (i.e., spawning output) and the terms spawning biomass and spawning output are used interchangeably through the report.

@tbl-es-sb

```{r, results = "asis"}
#| label: tbl-es-sb
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) ssb_es_cap 
#| tbl-pos: H

ssb_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  gt::cols_width(
    tidyselect::everything() ~ px(75)
  ) |>
  gt::as_latex()

```


```{r} 
#| label: fig-es-sb
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of spawning output for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts7_Spawning_output_with_95_intervals.png")
)

```


```{r} 
#| label: fig-es-depl
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fraction of unfished spawning output for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts9_Relative_spawning_output_intervals.png")
)

```

## Recruitment {-}

```{r, results = "asis"}
#| label: tbl-es-recr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recr_es_cap 
#| tbl-pos: H

recr_es_table |>
  dplyr::filter(Year < 2025) |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  gt::cols_width(
    tidyselect::everything() ~ px(75)
  ) |>
  gt::as_latex()

```

```{r} 
#| label: fig-es-recruits
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of age-0 recruits for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png")
)

```

```{r} 
#| label: fig-es-recdevs
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of recruitment deviations for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "recdevs2_withbars.png")
)

```


## Exploitation Status {-}


```{r, results = "asis"}
#| label: tbl-es-spr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) spr_es_cap 
#| tbl-pos: H

spr_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:7,
    decimals = 3
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  gt::cols_width(
    tidyselect::everything() ~ px(80)
  ) |>
  gt::as_latex()

```


```{r} 
#| label: fig-es-spr
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fishing intensity for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR3_ratiointerval.png")
)

```


## Ecosysystem Consideration {-}


## Reference Points {-}

```{r, results='asis'}
#| label: 'hcr-es'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "harvest_control_rule.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```

The estimated `r end_year` spawning biomass relative to unfished equilibrium spawning output is `r depl_final`, is below the management target of 40 percent of unfished spawning output. The fishing intensity is estimated to be below the current management harvest rate limit ($\text{SPR}_{45\%}$) at the end of the time series but is estimated to have been above between the 1970s to the late-2010s (@fig-es-spr). The interaction between the relative spawning output and the ratio of the estimated \gls{spr} to the management target ($\text{SPR}_{45\%}$) indicates that the stock is in the lower quadrent indicating that the overfishing is not occurring and that relative spawning output is below the management targets (@fig-es-spr-phase). The current estimate of fraction unfished is larger than the \gls{msy}, below the \gls{spr} target and relative target biomass reference points (@fig-es-yield) based on $h$ of `r h`.

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the most recent year of the model, `r end_year` (@tbl-ref-points-es). The estimated sustainable total yield inclusive of both landings and discards was `r catch_spr` mt according to $\text{SPR}_{45\%}$. The spawning output equivalent to 40 percent of unfished spawning biomass ($\text{SB}_{40\%}$) calculated using the SPR target, ($\text{SPR}_{45\%}$), was `r sb_spr` mt.

```{r, results = "asis"}
#| label: tbl-ref-points-es
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) reference_points_cap 
#| tbl-pos: H

reference_points_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    rows = c(1:4, 7, 10, 12, 15, 17, 20),
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = 2:4,
    rows = c(5, 8, 9, 13:14, 18:19),
    decimals = 3
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>  
  gt::tab_style(
    style = list(
      gt::cell_text(style = "italic")
    ),
    locations = gt::cells_body(
      columns = "Reference Point",
      rows = tidyselect::starts_with("Reference")
    )
  ) |>
  gt::sub_missing(
    columns = tidyselect::everything(),
    missing_text = "---"
  ) |>
  gt::as_latex()

```


```{r} 
#| label: fig-es-spr-phase
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Phase plot of fishing intensity versus fraction unfished for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR4_phase.png")
)

```


```{r} 
#| label: fig-es-yield
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated yield curve with reference points for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "yield2_yield_curve_with_refpoints.png")
)

```

## Management Performance {-}


```{r, results = "asis"}
#| label: tbl-es-management
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recent_management_cap 
#| tbl-pos: H

recent_management_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:6,
    decimals = 0
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  gt::as_latex()

```


## Evaluation of Scientific Uncertainty {-}


```{r, results='asis'}
#| label: 'scientific-uncertainty-es'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "scientific_uncertainty.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```




## Harvest Projections and Decision Tables {-}


```{r, results = "asis"}
#| label: tbl-es-gmt-catch
#| warning: false
#| echo: false
#| tbl-cap: "Removal assumptions in metric tons (mt) provided by the Groundfish Management Team for 2025-2026 by fleet."
#| tbl-pos: H

gmt_catch |>
  dplyr::select(-comment, -Seas) |>
  dplyr::mutate(
    Fleet = dplyr::case_when(
      Fleet == 1 ~ "Trawl",
      Fleet == 2 ~ "HKL",
      Fleet == 3 ~ "Pot",
      Fleet == 4 ~ "Trawl Discard",
      Fleet == 5 ~ "HKL Discard",
      Fleet == 6 ~ "Pot Discard"
    )
  ) |>
  tidyr::pivot_wider(
    names_from = Fleet,
    values_from = `dead(B)`
  ) |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(2:7),
    decimals = 1
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::as_latex()

```

::: {.landscape}

```{r, results = "asis"}
#| label: tbl-es-projections
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) projections_cap 
#| tbl-pos: H

# Note to self: the ABC for 2027 and 2028 is not provided due to the HCR
# need to add those in to the table.

projections_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(2:5, 7:9),
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = c(6, 10),
    decimals = 3
  ) |>
  gt::tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::sub_missing(
    columns = tidyselect::everything(),
    missing_text = "---"
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  gt::cols_width(
    everything() ~ px(75)
  ) |>
  gt::as_latex()

```

:::

## Unresolved Problems and Major Uncertainties {-}


```{r, results='asis'}
#| label: 'uncertainties-es'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "unresolved_uncertainties.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```

## Research and Data Needs {-}


```{r, results='asis'}
#| label: 'research-data-es'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "research_data_needs.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```

## Risk Table {-}

ADD RISK TABLE HERE

```{r, results='asis'}
#| label: 'risk-table'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "risk_table.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```
