
## Executive Summary {-}

### Stock Description {-}

This assessment reports the status of the `r spp` (*Anoplopoma fimbria*, or `black cod') resource off the coast of the United States (U.S.) from southern California to the U.S.-Canadian border using data through `r data_end_year`. The resource is modeled as a single stock, however `r spp` do disperse to and from offshore sea mounts and along the coastal waters of the continental U.S., Canada, and Alaska and across the Aleutian Islands to the western Pacific. Their movement is not explicitly accounted for in this analysis.

### Catches {-}

Check to see if referencing works @tbl-es-catches


```{r, results = "asis"}
#| label: tbl-es-catches
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) catches_es_cap 
#| tbl-pos: H

catches_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:6,
    decimals = 0
  ) |>
  tab_options(
    table.font.size = 12,
    # if want to not have tables the width of the page include the below option:
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  #cols_width(
  #  everything() ~ px(75)
  #) |>
  as_latex()

```

```{r} 
#| label: fig-es-landings
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Landings in metric tons (mt) by year for trawl, pot, and hook-and line gears."
#| fig-alt: "Landings before the 1970s were less than 5,000 mt by year with landings sharply increasing in the 1970s and early 1980s and landings since ranging between approximately 5,000 to 7,500 mt by year."

knitr::include_graphics(
  file.path(model_dir, "plots", "catch2_landings_stacked.png")
)

```

### Data and Assessments {-}


### Stock Output and Dynamics {-}

@tbl-es-sb

```{r, results = "asis"}
#| label: tbl-es-sb
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) ssb_es_cap 
#| tbl-pos: H

ssb_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```


```{r} 
#| label: fig-es-sb
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of spawning biomass for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts7_Spawning_output_with_95_intervals.png")
)

```


```{r} 
#| label: fig-es-depl
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fraction of unfished spawning biomass for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts9_Relative_spawning_biomass_intervals.png")
)

```

### Recruitment {-}

```{r, results = "asis"}
#| label: tbl-es-recr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recr_es_cap 
#| tbl-pos: H

recr_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:4,
    decimals = 0
  ) |>
  fmt_number(
    columns = 5:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```

```{r} 
#| label: fig-es-recruits
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of age-0 recruits for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png")
)

```

```{r} 
#| label: fig-es-recdevs
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of recruitment deviations for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "recdevs2_withbars.png")
)

```


### Exploitation Status {-}


```{r, results = "asis"}
#| label: tbl-es-spr
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) spr_es_cap 
#| tbl-pos: H

spr_es_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:7,
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(80)
  ) |>
  as_latex()

```


```{r} 
#| label: fig-es-spr
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fishing intensity for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR3_ratiointerval.png")
)

```


### Ecosysystem Consideration {-}


### Reference Points {-}


```{r, results = "asis"}
#| label: tbl-ref-points
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) reference_points_cap 
#| tbl-pos: H

reference_points_table |>
  gt() |>
  fmt_number(
    columns = 2:4,
    rows = c(1:4, 7, 10, 12, 15, 17, 20),
    decimals = 0
  ) |>
  fmt_number(
    columns = 2:4,
    rows = c(5, 8, 9, 13:14, 18:19),
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>  
  tab_style(
    style = list(
      cell_text(style = "italic")
    ),
    locations = cells_body(
      columns = "Reference Point",
      rows = starts_with("Reference")
    )
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "---"
  ) |>
  #cols_width(
  #  starts_with("R") ~ px(300),
  #  starts_with("E") ~ px(65),
  #  ends_with("l") ~ px(65)
  #) |>
  as_latex()

```

```{r} 
#| label: fig-es-spr-phase
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Phase plot of fishing intensity versus fraction unfished for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR4_phase.png")
)

```



```{r} 
#| label: fig-es-yield
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated yield curve with reference points for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "yield2_yield_curve_with_refpoints.png")
)

```

### Management Performance {-}


```{r, results = "asis"}
#| label: tbl-es-management
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) recent_management_cap 
#| tbl-pos: H

recent_management_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = 2:6,
    decimals = 0
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  #cols_width(
  #  everything() ~ px(75)
  #) |>
  as_latex()

```


### Evaluation of Scientific Uncertainty {-}


### Harvest Projections and Decision Tables {-}

\blandscape

```{r, results = "asis"}
#| label: tbl-es-projections
#| warning: false
#| echo: false
#| eval: !expr eval_tables 
#| tbl-cap: !expr if(eval_tables) projections_cap 
#| tbl-pos: H

projections_table |>
  gt::gt() |>
  gt::fmt_number(
    columns = c(2:5, 7:9),
    decimals = 0
  ) |>
  gt::fmt_number(
    columns = c(6, 10),
    decimals = 3
  ) |>
  tab_options(
    table.font.size = 12,
    latex.use_longtable = TRUE 
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "---"
  ) |>
  gt::cols_align(
    align = "center"
  ) |>
  cols_width(
    everything() ~ px(75)
  ) |>
  as_latex()

```

\elandscape

### Unresolved Problems and Major Uncertainties {-}

### Research and Data Needs {-}

