# Management


## Reference Points

```{r, results='asis'}
#| label: 'hcr'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "harvest_control_rule.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


The derived stock status in `r end_year` is `r depl_final*100` percent, below the management target of 40 percent. The fishing intensity is estimated to be below the current management harvest rate limit ($\text{SPR}_{45\%}$) at the end of the time series but is estimated to have been above between the 1970s to the late-2010s (@fig-spr). The interaction between the stock status and the ratio of the estimated \gls{spr} to the management target ($\text{SPR}_{45\%}$) indicates that the stock is in the lower quadrant indicating that the overfishing is not occurring and that stock status is below the management target (@fig-spr-phase). The current estimate of stock status is larger than the \gls{msy}, below the \gls{spr} target and stock status reference points (@fig-yield) based on $h$ of `r h`.

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the final year of the model, `r end_year` (@tbl-ref-points). The estimated sustainable total yield inclusive of both landings and discards was `r catch_spr` mt according to $\text{SPR}_{45\%}$. The spawning output equivalent to 40 percent of unfished spawning output ($\text{SB}_{40\%}$) calculated using the SPR target, ($\text{SPR}_{45\%}$), was `r sb_spr` mt.

## Evaluation of Scientific Uncertainty


```{r, results='asis'}
#| label: 'scientific-uncertainty'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "scientific_uncertainty.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```

## Harvest Projections and Decision Tables

The projection of spawning ouput, stock status, and harvest limits were developed using the base model. The total catches in `r end_year` and `r end_year + 1` were fixed at `r fixed_catch[1]` and `r fixed_catch[2]` mt, respectively, based on recommendations from the \gls{gmt} (@tbl-gmt-catch). The \gls{abc} values were projected assuming a category 1 time-varying sigma with an initial value of 0.50 in 2027 and a management risk tolerance (P$^*$) of 0.45.  

The catches during the projection period, `r end_year + 2`--`r end_year + 12` were set equal to the year-specific \glsplural{abc} (@tbl-projections). The stock size and status increase during the projection due to the estimated large recruitments in 2020 and 2021 maturing and entering the spawning population, resulting in future \glsplural{ofl} and \glsplural{abc}.


## Regional Management Considerations

`r Spp` is currently modeled as a coastwide stock with corresponding coastwide \gls{ofl} and \gls{abc}. The coastwide \gls{abc} is then apportioned into two area-specific \glsplural{acl} north and south of 36$^\circ$ N. latitude. For the 2019 assessment, the \gls{pfmc} adopted a methodology to split the \gls{acl} using a five-year rolling average biomass estimate by area from the \gls{s-wcgbt} ([November 2019 \gls{pfmc} Meeting Decision Document](https://www.pcouncil.org/documents/2019/11/november-2019-decision-document.pdf/)). This historical management line corresponds with a recent data-driven analysis of `r spp` growth that suggests a difference in growth rates north and south of $36^\circ$ N. latitude [@kapur_oceanographic_2020]. The estimates represent the relative distribution of the `r spp` population observed by the survey, not the entire population. The coastwide \gls{ofl} and \gls{abc} values and the area based \gls{acl} are shown in @tbl-management.

The estimated proportion of the `r data_end_year - 6`--`r data_end_year` observed biomass by the \gls{s-wcgbt} is shown in @tbl-prop-area.  

## Risk Table {#risk}

Oceanographers and stock assessment scientists jointly constructed a risk table to identify the main ecosystem and environmental processes for `r spp` that should be considered in conjunction with available data used in the stock assessment and the structural uncertainty embedded in the assessment (@tbl-risk-table). The group considered ecosystem and environmental indicators for `r spp` recruitment, habitat, prey, predators, competitors, and the recent climate vulnerability assessment [@McClure:2023:VCC]. Overall, the group considered these conditions to generally be favorable (Level 1), despite unresolved uncertainty with linkages between trends in primary predators and competitors and a overall climate vulnerability score of *moderate*.
Data inputs for `r spp` were scored as favorable (Level 1) given confidence in historical catch reconstructions, a comprehensive and representative survey index of abundance (WCGBTS), well sampled age data beginning at age zero, an informative recruitment index, and information on time- (growth) and -spatially-varying (maturity) life history characteristics across the stock range. The ability for the stock assessment model to fit the data and capture key sources of structural uncertainty was also scored as favorable (Level 1). Despite the need to fix certain productivity parameters (e.g., steepness and recruitment variability) or simplify some assumptions (e.g., sex invariant *M* and no net loss from emigration or immigration), model complexity seemed to balance well potential bias and variance in a way that stabilized `r spp` population dynamics across a wide breadth of alternative data and model structures (see section \ref{sensitivities} for sensitivity evaluations).

## Unresolved Problems and Major Uncertainties


```{r, results='asis'}
#| label: 'uncertainties'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "unresolved_uncertainties.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


## Research and Data Needs


```{r, results='asis'}
#| label: 'research-data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "research_data_needs.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


