# Management


## Reference Points

```{r, results='asis'}
#| label: 'hcr'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "harvest_control_rule.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


The estimated `r end_year` spawning biomass relative to unfished equilibrium spawning output is `r depl_final`, is below the management target of 40 percent of unfished spawning output. The fishing intensity is estimated to be below the current management harvest rate limit ($\text{SPR}_{45\%}$) at the end of the time series but is estimated to have been above between the 1970s to the late-2010s (@fig-spr). The interaction between the relative spawning output and the ratio of the estimated \gls{spr} to the management target ($\text{SPR}_{45\%}$) indicates that the stock is in the lower quadrent indicating that the overfishing is not occurring and that relative spawning output is below the management targets (@fig-spr-phase). The current estimate of fraction unfished is larger than the \gls{msy}, below the \gls{spr} target and relative target biomass reference points (@fig-yield) based on $h$ of `r h`.

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the most recent year of the model, `r end_year` (@tbl-ref-points). The estimated sustainable total yield inclusive of both landings and discards was `r catch_spr` mt according to $\text{SPR}_{45\%}$. The spawning output equivalent to 40 percent of unfished spawning biomass ($\text{SB}_{40\%}$) calculated using the SPR target, ($\text{SPR}_{45\%}$), was `r sb_spr` mt.


## Evaluation of Scientific Uncertainty


```{r, results='asis'}
#| label: 'scientific-uncertainty'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "scientific_uncertainty.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


## Harvest Projections and Decision Tables

The `r end_year` assessment for `r spp` off the U.S. West Coast was assigned a category 1 determination by the \gls{ssc} to the \gls{pfmc} (UPDATE TEXT AS NECESSARY). The projection of spawning biomass, fraction unfished, and harvest limits was developed using the base model. The total catches in `r end_year` and `r end_year + 1` were set at `r fixed_catch[1]` and `r fixed_catch[2]` mt, respectively, based on recommendations from the \gls{gmt}. The \gls{abc} values were estimated using a category 1 time-varying $\sigma_y$ starting at 0.50 combined with a P$^{*}$ value of 0.45. 

@tbl-gmt-catch

The catches during the projection period, `r end_year + 2`--`r end_year + 12` were set equal to the year-specific \glsplural{abc} (@tbl-projections). The spawning biomass and fraction unfished increase during the projection due to the estimated large recruitments in 2020 and 2021 maturing and entering the spawning population, resulting in future \glsplural{ofl} and \glsplural{abc}.


## Regional Management Considerations

`r Spp` is currently modeled as a coastwide stock with corresponding coastwide \gls{ofl}, \gls{abc}, and \gls{acl} values. The coastwide \gls{acl} is then apportioned into two area-specific \glsplural{acl} north and south of 36$^\circ$ N. latitude. For the 2019 assessment, the \gls{pfmc} adopted a methodology to split the \gls{acl} using a five-year rolling average biomass estimate by area from the \gls{s-wcgbt} ([November 2019 \gls{pfmc} Meeting Decision Document](https://www.pcouncil.org/documents/2019/11/november-2019-decision-document.pdf/)). This historical management line corresponds with a recent data-driven analysis of `r spp` growth that suggests a difference in growth rates north and south of $36^\circ$ N. latitude [@kapur_oceanographic_2020]. The estimates represent the relative distribution of the `r spp` population observed by the survey, not the entire population. Additionally, it is likely that fish from more northerly regions are migrating into U.S. West Coast waters (pers. comm., L. Rogers), which may bias the survey estimates of the distribution of fish in each region. Thus, these results should be interpreted with caution. The coastwide \gls{ofl} and \gls{abc} values and the area based \gls{acl} are shown in @tbl-area-spex.

The estimated proportion of the `r data_end_year - 6`--`r data_end_year` observed biomass by the \gls{s-wcgbt} is shown in @tbl-prop-area.  


## Unresolved Problems and Major Uncertainties


```{r, results='asis'}
#| label: 'uncertainties'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "unresolved_uncertainties.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


## Research and Data Needs


```{r, results='asis'}
#| label: 'research-data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "research_data_needs.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


