# Management


## Reference Points

```{r, results='asis'}
#| label: 'hcr'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "harvest_control_rule.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


The derived stock status in `r end_year` is `r depl_final*100` percent, below the management target of 40 percent. The fishing intensity is estimated to be below the current management harvest rate limit ($\text{SPR}_{45\%}$) at the end of the time series but is estimated to have been above it between the 1970s to the late-2010s (@fig-spr). The interaction between the stock status and the ratio of the estimated \gls{spr} to the management target ($\text{SPR}_{45\%}$) indicates that the stock is in the lower quadrant indicating that the overfishing is not occurring and that stock status is below the management target (@fig-spr-phase). The current estimate of stock status is larger than the \gls{msy}, below the \gls{spr} target and stock status reference points (@fig-yield) based on $h$ of `r h`.

Reference points were calculated using the estimated selectivities and catch distributions among fleets in the final year of the model, `r end_year` (@tbl-ref-points). The estimated sustainable total yield inclusive of both landings and discards was `r catch_spr` mt according to $\text{SPR}_{45\%}$. The spawning output equivalent to 40 percent of unfished spawning output ($\text{SB}_{40\%}$) calculated using the SPR target, ($\text{SPR}_{45\%}$), was `r sb_spr` mt.

## Management Performance 

Sablefish management includes a rich history of seasons, size-limits, trip-limits, and a complex permit system. `r Spp` is currently modeled as a coastwide stock with corresponding coastwide \gls{ofl} and \gls{abc}. The coastwide \gls{abc} is then apportioned into two area-specific \glsplural{acl} north and south of 36$^\circ$ N. latitude. This historical management line corresponds with a recent data-driven analysis of `r spp` growth that suggests a difference in growth rates north and south of $36^\circ$ N. latitude [@kapur_oceanographic_2020]. The coastwide \gls{ofl} and \gls{abc} values and the area-based \gls{acl} are shown in @tbl-management.

For the 2019 assessment, the \gls{pfmc} adopted a methodology to determine the area-based \gls{acl} using a five-year rolling average relative biomass estimate from the WCGBTS ([November 2019 \gls{pfmc} Meeting Decision Document](https://www.pcouncil.org/documents/2019/11/november-2019-decision-document.pdf/)). The relative biomass estimates represent the distribution of the `r spp` population observed by the WCGBTS, not the entire population.  

The proportion of biomass observed by the WCGBTS north and south of 36$^\circ$ N. latitude for the most recent five years is shown in @tbl-prop-area.  


## Evaluation of Scientific Uncertainty


```{r, results='asis'}
#| label: 'scientific-uncertainty'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "scientific_uncertainty.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```

## Harvest Projections and Decision Tables {#forecast}

The projection of spawning ouput, stock status, and harvest limits were developed using the base model. The total catches in `r end_year` and `r end_year + 1` were fixed at `r fixed_catch[1]` and `r fixed_catch[2]` mt, respectively, based on recommendations from the \gls{gmt} (@tbl-gmt-catch). The \gls{abc} values were projected assuming a category 1 time-varying sigma with an initial value of 0.50 in 2027 and a management risk tolerance (P$^*$) of 0.45.  

The catches during the projection period, `r end_year + 2`--`r end_year + 12` were set equal to the year-specific \glsplural{abc} (@tbl-projections). The stock size and status increase during the projection due to the estimated large recruitments in 2020 and 2021 maturing and entering the spawning population, resulting in future \glsplural{ofl} and \glsplural{abc}.

Two alternative catch streams based upon two options for P$^*$, 0.45 and 0.40, were created for the decision table (@tbl-decision-table). Each alternative catch stream assumes a category 1 time varying sigma with an initial value of 0.50 in 2027.  The axes of uncertainty in the decision table are based on the uncertainty around natural mortality. The low state of nature assumed a higher natural mortality of 0.095 yr$^{-1}$ and the high state of nature assumed a lower natural mortality of 0.070 yr$^{-1}$.  The low and high state of nature were run in a deterministic manner where they assumed the same parameter values, including annual recruitment deviations, as the base model except for natural mortality.  

## Regional Management Considerations

`r Spp` is currently modeled as a coastwide stock with a corresponding coastwide \gls{ofl} and \gls{abc}. The coastwide \gls{abc} is then apportioned into two area-specific \glsplural{acl} north and south of 36$^\circ$ N. latitude. This historical management line corresponds with a recent data-driven analysis of `r spp` growth that suggests a difference in growth rates north and south of $36^\circ$ N. latitude [@kapur_oceanographic_2020]. The coastwide \gls{ofl} and \gls{abc} values and the area-based \gls{acl} are shown in @tbl-management.

For the 2019 assessment, the \gls{pfmc} adopted a methodology to determine the area-based \gls{acl} using a five-year rolling average biomass estimate by area from the \gls{s-wcgbt} ([November 2019 \gls{pfmc} Meeting Decision Document](https://www.pcouncil.org/documents/2019/11/november-2019-decision-document.pdf/)).  Estimates presented are based on the relative distribution of the `r spp` population observed by the survey, not the entire population. The estimated proportion of the `r data_end_year - 6`--`r data_end_year` observed biomass by the \gls{s-wcgbt} is shown in @tbl-prop-area.  

## Risk Table {#risk}

Oceanographers and stock assessment scientists jointly constructed a risk table to identify the main ecosystem and environmental processes for `r spp` that should be considered in conjunction with available data used in the stock assessment and the structural uncertainty embedded in the assessment (@tbl-risk-table). The group considered ecosystem and environmental indicators for `r spp` recruitment, habitat, prey, predators, competitors, and the recent climate vulnerability assessment [@McClure:2023:VCC]. Overall, the group considered these conditions to generally be neutral (Level 2) when considering, among other things, unresolved uncertainty with linkages between trends in primary predators and competitors and a overall climate vulnerability score of moderate.
Data inputs for `r spp` were scored as favorable (Level 1) given confidence in historical catch reconstructions, a comprehensive and representative survey index of abundance (WCGBTS), well sampled age data beginning at age zero, an informative recruitment index, and information on time- (growth) and -spatially-varying (maturity) life history characteristics across the stock range. The ability for the stock assessment model to fit the data and capture key sources of structural uncertainty was also scored as favorable (Level 1). Despite the need to fix certain productivity parameters (e.g., steepness and recruitment variability) or simplify some assumptions (e.g., sex invariant *M* and no net loss from emigration or immigration), model complexity seemed to balance well potential bias and variance in a way that stabilized `r spp` population dynamics across a wide breadth of alternative data and model structures (see Section \ref{sensitivities} for sensitivity evaluations).

To understand the impact of ecosystem and environmental conditions on sablefish, we evaluated recent trends in drivers of recruitment, predators, and prey, along with context from the climate vulnerability assessment (CVA) in @McClure:2023:VCC. We did not consider habitat or non-fisheries human activities (such as offshore wind development) during this evaluation. In considering predators and prey effects here, we focus on recent trends over the last five years; we assume that long term trends are implicitly represented in the stock assessment and do not represent notable changes that warrant inclusion in the risk table. 

### Recruitment

Environmental drivers of recruitment for sablefish have received attention in the literature and prior stock assessments [@schirripa2006; @schirripa2009icesjms;  @tolimieri2018oceanographic; @haltuch_status_2019; @johnson_status_2023; @tolimieri2023sea]. Prior assessments have examined or included sea-level in the northern portion of the West Coast as an environmental index of recruitment in the stock assessment model [@haltuch_status_2019; @johnson_status_2023]. More recent work examining the utility of output from the GLORYS model suggests that cross-shelf transport during the pelagic larval and benthic juvenile stages is a good predictor of sablefish recruitment (see Section \ref{rec-appendix}). 

A species distribution model (`sdmTMB`, delta-lognormal, spatial and spatiotemporal autocorrelation, scaled depth as a quadratic variable) used the WCGBTS data to estimate the distribution and abundance of  sablefish juveniles (29 cm total length, usually age-0 but may include age-1 fishes in years following high recruitment). Prior to 2014, juveniles were more frequent and more abundant south of Cape Mendocino. However, from 2014--2019 (during the strong marine heatwave years) juveniles were more abundant in the north. From 2019 juvenile abundance has been relatively evenly distributed across the West Coast. 

Data from the WCGBTS and predictions from the northern sea-level index and the GLORYS index indicate good recruitment (juvenile abundance) conditions in recent years (with the exception of 2022) with some exceptionally high recruitment years. Data and projections for 2025 are not available. Upon a comprehensive risk table examination, recruiment was not explicitly incorporated into the overall rating for ecosystem and environmental conditions because of the explicit inclusion of the recruitment index in the assessment model.

### Distribution and Habitat Considerations

The distribution of sablefish biomass was examined using data from the WCGBTS and species distribution modeling using `sdmTMB`. The delta-lognormal model included spatial and spatiotemporal autoregression including scaled depth as a smoothed parameter (3 knots). The center of gravity of sablefish biomass has shifted north by approximately 2$^\circ$ N. since 2013 to 41$^\circ$ 5$\prime$ N. latitude in 2024 but only by approximately 1$^\circ$ N. since the start of the WCGBTS in 2003. As a result, sablefish have become more available to northern ports, especially Astoria. The shift north coincides with stronger recruitment (juvenile abundance) north of Cape Mendocino concurrent with the large marine heatwave of the 2014--2016 period.

###  Predators and Prey and Trophic Overlap with Competitors

We did not see evidence for recent changes in potential predation on sablefish. California sea lions, sperm whales, porpoises, skates, and fur seals were estimated to account for 97 percent of predation mortality on sablefish (Ecopath food web model, @field_top_2006, @koehn_developing_2016, recently revised by C. Best-Otubu, P.Y. Hernvann, N. Lezama Ochoa, I. Kaplan, with 2006 as the base year for calculation of food web interactions and mass balance). While California sea lions have shown substantial recovery since the 1970s [@carretta2022us], the population appears to have stabilized around 2010. More recently, the 2023 sea lion pup cohort (see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) Figure N.1) exhibited low growth rate. Assessments for sperm whales [@carretta2022us] and Dall’s porpoises [@carretta2022us] do not suggest strong abundance trends for 1991--2018 and do not provide more recent data. Assessments for big skate [@taylor2019status] and longnose skate [@gertseva2019stock] estimated stable population sizes for both species through 2019. Fur seal populations also have recovered from low levels, but the stock assessment report includes population estimates only to 2013. 

Available evidence suggests average or adequate prey availability. Common prey items for adult sablefish are: deep small rockfish (including longspine thornyhead), Pacific hake, Pacific herring, cephalopods, deposit feeders (including amphipods, isopods, small crustacea), deep demersal fish (grenadier, eelpout), and carrion [@wippel]. Juvenile diets include: small planktivorous fish, euphausiids, gelatinous zooplankton, and deposit feeders [@wippel].

Euphausiid abundance appears to be stable or slightly increasing in Central California over the last five years (RREAS survey, see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) Figure I.2) with variable abundance in Northern waters between 42$^\circ$-46$^\circ$ N. latitude but with no strong trend and the mean of the last five years near the long term average (data from B. Wells, NOAA SWFSC Pers. Comm.). While, krill length and biomass indices from the Trinidad Head survey line were near average for most of 2024 (see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) Figure 3.2), the coastwide krill abundance index, derived from acoustic data was 2nd lowest in 2023 (no 2024 data available) since the beginning  of sampling in 2007 (see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2024/02/agenda-item-h-1-a-cciea-team-report-1-2023-2024-california-current-ecosystem-status-report-electronic-only.pdf/) Figure 3.3). Production of fledgling Cassin’s auklet, which feeds on krill, was near average at Southeast Farallon Island suggesting typical feeding conditions (see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) Figure M.2).

Production of juvenile Pacific hake was in above average in 2020 and 2021 [@JTC2025]. Though there is high uncertainty in the stock assessment estimates of recruitment for more recent years [@JTC2025], field surveys provide evidence of continued good Pacific hake recruitment in recent years (see [California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) Figure I2). Herring abundance sampled by JSOES in the Northern California Current has been stable in recent years, though the 2023 acoustic survey estimates of coastwide herring abundance suggest a decline from 2019 levels [@stierhoff2024distribution]. The longspine thornyhead stock was estimated to be at healthy levels of abundance in 2013 [@stephens2013stock], but we cannot infer the recent (approximately 2021--2025) status of the stock. 

#### Trophic Overlap

Euphausiids are a substantial component of the sablefish diet. Potential competitors for euphausiids were identified from the Ecopath food web model described above. Carnivorous zooplankton, mesopelagic fish, adult Pacific hake, large jellyfish, other cephalopods, sardine, and juvenile Pacific hake are estimated to account for 75 percent of total predation mortality on euphausiids, with adult and juvenile Pacific hake together accounting for 16 percent of predation on euphausiids. There is substantial uncertainty in these estimates, especially because of high uncertainty regarding abundances of these invertebrates and mesopelagic fish. In comparison, sablefish account for less than 1 percent (0.1 percent) of predation mortality on euphausiids. Though we cannot quantitatively determine that euphausiid abundance is limiting for sablefish, there is some potential for competition by Pacific hake and sablefish for prey (euphausiids), given 1) some spatial overlap of sablefish and Pacific hake on the outer continental shelf near the shelf break, and 2) based on the 2025 assessment of Pacific hake, Pacific hake total biomass in 2025 is 72 percent higher in 2025 than in 2006, which was the base year for the Ecopath calculations. In contrast, though lonspine thornyhead appear common in sablefish diets, there appears to be less competition for longspine thornyhead by other predators: 84 percent of the predation pressure on longspine thornyhead is estimated in the Ecopath model to derive from sablefish. 

###  Climate Vulnerability Assessment Rank

@McClure:2023:VCC found that sablefish had a climate vulnerability of moderate, based on consideration of both exposure and sensitivity. Climate exposure was high, due largely to potential impacts on these species (and many others) from ocean acidification and mean sea surface temperature. These exposure metrics in the CVA represent expected change between a historical period (1955--2005) and a future period (2006--2055). Biological sensitivity was ranked moderate, due to slow population growth rate and early life history survival and settlement requirements. Relevant to these Risk Tables and recent trends, the overall context from the rapid, expert-opinion CVA for this species suggests particular attention should be paid to any recent changes in ocean acidification (related to potential effects on prey such as zooplankton) or in environmental conditions relevant to early life history and recruitment. We consider the effects of prey availability and recruitment to be well informed in other sections of this risk table and the assessment, and as a result the CVA ranking was not included in our final scoring of the ecosystem and environmental considerations.

### Theoretical Projections

Accounting for each factor considered in the risk table (@tbl-risk-table), the overall conditions for sablefish is currently considered favorable which could support the application of scientific uncertainty ($\sigma$) lower than the default ($\sigma$ of 0.50) for a category one assessment for projecting future harvest specifications. @tbl-risk-table-projections shows an theoretical projection using a $\sigma$ of 0.25 and a P* of 0.45 that could be considered if the use of risk tables is adopted by management for future stock assessments and a $\sigma$ of 0.25 is deemed appropriate by the Scientific and Statistical Committee.


## Unresolved Problems and Major Uncertainties


```{r, results='asis'}
#| label: 'uncertainties'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "unresolved_uncertainties.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


## Research and Data Needs


```{r, results='asis'}
#| label: 'research-data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child(here::here("report", "shared_text", "research_data_needs.qmd"), quiet = TRUE)
cat(a, sep = '\n')
```


