# Figures 
 
```{r} 
#| label: 'set-rda-dir-figs'
#| echo: false 
#| warning: false 
#| eval: true 
rda_dir <- here::here('rda_files')
``` 


## Data


```{r} 
#| label: fig-map
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Map of the assessment area."
#| fig-alt: "Map of the U.S. West Coast with a line indicating that this assessment covers the full U.S. West Coast."
#| fig-width: 7
#| fig-height: 8

mod.names <- c("Management  Range", "Assessment Range")
mod.cols  <- c("blue","lightblue", "yellow")

map('worldHires', regions=c("Canada","Mexico"),
    xlim=c(-130, -114), ylim=c(31, 51),
    col='grey', fill=TRUE, interior=TRUE, , lwd=1)

# map with US states
map('state', regions=c("Wash","Oreg","Calif","Idaho", "Montana","Nevada","Arizona","Utah"),
    add = TRUE,
    col = 'grey', fill = TRUE, interior = TRUE, lwd = 1)
axis(2, at = seq(32, 50, 2), lab = paste0(seq(32, 50, 2), "°N"), las=1)
axis(1, at = seq(-130, -114, 4), lab = paste0(abs(seq(-130, -114, 4)), "°W"))

## add vertical lines indicating range for each stock
# North of 36
text(-128-.8, 40.5, mod.names[1], srt=90)
latrange <- c(32.5, 36) + c(.2, -.2)
lines(rep(-128, 2), latrange, lwd=10, col=mod.cols[1])

# South of 36
latrange <- c(36, 48.5) + c(.2, -.2)
lines(rep(-128, 2), latrange, lwd = 10, col=mod.cols[2])
#text(-128-.8, mean(latrange), mod.names[1], srt=90)
# Assessment area
latrange <- c(32.5, 48.5) + c(.2, -.2)
lines(rep(-126,2), latrange, lwd=10, col=mod.cols[3])
text(-126-.8, mean(latrange), mod.names[2], srt=90)

text(-120, 50, "Canada")
text(-120, 47, "Washington")
text(-121, 44, "Oregon")
text(-119.5, 37, "California")
text(-115.5, 32.1, "Mexico")


``` 



```{r} 
#| label: fig-data
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Data used in the base model."
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "data_plot.png")
)

``` 

\newpage

```{r} 
#| label: fig-landings
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Landings in metric tons (mt) by year for trawl, pot, and hook-and line gears."
#| fig-alt: "Landings before the 1970s were less than 5,000 mt by year with landings sharply increasing in the 1970s and early 1980s and landings since ranging between approximately 5,000 to 7,500 mt by year."

knitr::include_graphics(
  file.path(model_dir, "plots", "catch2_landings_stacked.png")
)

```


\pagebreak

```{r} 
#| label: fig-data-trawl-1
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt1mkt0_page1.png")
)

```

\pagebreak

```{r} 
#| label: fig-data-trawl-2
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt1mkt0_page2.png")
)

```

\pagebreak

```{r} 
#| label: fig-data-hkl-1
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt2mkt0_page1.png")
)

```

\pagebreak

```{r} 
#| label: fig-data-hkl-2
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt2mkt0_page2.png")
)

```

\pagebreak

```{r} 
#| label: fig-data-pot-1
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt3mkt0_page1.png")
)

```

\pagebreak

```{r} 
#| label: fig-data-pot-2
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: ""
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "comp_agedat_flt3mkt0_page2.png")
)

```


\pagebreak

Add catch comparison figure

\pagebreak


Add discard length and age figures

\pagebreak


```{r} 
#| label: fig-index-stand
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Standardized indices overlaid. Each index is rescaled to have mean observation = 1.0."
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "index9_standcpueall.png")
)

```

\pagebreak

```{r} 
#| label: fig-qq-wcgbt
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "QQ plot for the NWFSC WCGBT survey."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "wcgbt_qq.png")
)

```

\pagebreak


```{r} 
#| label: fig-index-wcgbt
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Relative index of abundance for the NWFSC WCGBT survey."
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "index1_cpuedata_WCGBT.png")
)

```

\pagebreak


```{r} 
#| label: fig-positive-lat-wcgbt
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Proportion of tows where sablefish was present or absent across all years by latitude for the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-alt: ""

 nwfscSurvey::plot_proportion(
    data = wcgbt_catch |> 
      dplyr::mutate(new = factor(cpue_kg_km2 <= 0, levels = c(FALSE, TRUE), labels = c("Present", "Absent"))),
    column_factor = new,
    column_bin = Latitude_dd,
    width = 1,
    boundary = 0,
    bar_width = "equal"
  ) 

```


```{r} 
#| label: fig-positive-depth-wcgbt
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Proportion of tows where sablefish was present or absent across all years by depth (m) for the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-alt: ""

 nwfscSurvey::plot_proportion(
    data = wcgbt_catch |> 
      dplyr::mutate(new = factor(cpue_kg_km2 <= 0, levels = c(FALSE, TRUE), labels = c("Present", "Absent"))),
    column_factor = new,
    column_bin = Depth_m,
    width = 100,
    boundary = 0,
    bar_width = "equal"
  ) 

```

\pagebreak


```{r} 
#| label: fig-sex-ratio-wcgbt
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Sex ratio across all years by depth (m) for the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-alt: ""

 nwfscSurvey::plot_proportion(
    data = wcgbt_bio |> dplyr::filter(Sex != "U"),
    column_factor = Sex,
    column_bin = Depth_m,
    width = 100,
    boundary = 0,
    bar_width = "equal"
  ) 

```

\pagebreak

```{r} 
#| label: fig-wcgbt-map-density
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Sampling locations for sablefish collected by the NWFSC West Coast Groundfish Bottom Trawl survey between 2003 and 2024. Standard lengths (rounded down to the nearest cm) are shown on the left and age estimates (yr) are shown on the right. Lighter colors represent smaller or younger individuals and darker colors represent larger or older individuals. Latitudinal and inshore-offshore patterns were similar for males and females, thus sexes were combined."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "wcgbt_length_age_density.png")
)

```

\pagebreak

```{r} 
#| label: fig-wcgbt-depth-len-age
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Length (top; cm) and age (bottom; yr) compositions for sablefish collected by the NWFSC West Coast Groundfish Bottom Trawl Survey between 2003 and 2024 by sex and depth (m). Lighter colors represent smaller or younger individuals and darker colors represent larger or older individuals."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "wcgbts_length_age_compositions.png")
)

```

\pagebreak


```{r} 
#| label: fig-qq-nwfslope
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "QQ plot for the NWFSC Slope survey."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "nwfsc_slope_qq.png")
)

```

\pagebreak


```{r} 
#| label: fig-index-nwslope
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Relative index of abundance for the NWFSC Slope survey."
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "index1_cpuedata_NWFSC_Slope.png")
)

```

\pagebreak


```{r} 
#| label: fig-qq-afslope
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "QQ plot for the AFSC Slope survey."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "afsc_slope_qq.png")
)

```

\pagebreak

```{r} 
#| label: fig-index-afslope
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Relative index of abundance for the AFSC Slope survey."
#| fig-alt: ""

#knitr::include_graphics(
#  file.path(model_dir, "plots", "index1_cpuedata_AFSCSlope.png")
#)

```

\pagebreak


```{r} 
#| label: fig-qq-tri-early
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "QQ plot for the Triennial survey for the early period, 1980-1992."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "tri_early_qq.png")
)

```

\pagebreak



```{r} 
#| label: fig-qq-tri-late
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "QQ plot for the Triennial survey for the late period, 1995-2004."
#| fig-alt: ""

knitr::include_graphics(
  here::here("report", "figures", "tri_late_qq.png")
)

```

\pagebreak


```{r} 
#| label: fig-index-tri
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Relative index of abundance for the Triennial survey."
#| fig-alt: ""

knitr::include_graphics(
  file.path(model_dir, "plots", "index1_cpuedata_Triennial.png")
)

```

\pagebreak


## Biology

```{r} 
#| label: fig-maturity-data
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Samples by age that were determined to be functionally mature and immature north and south of 36 degrees north latitude. Immature fish have a functional maturity of 0 and mature fish have a functional maturity of 1. Blue vertical dashed line at age 7."
#| fig-alt: "Age associated with immature and mature fish by area."

load(here::here("data", "maturity_data.rda"))

ggplot2::ggplot(maturity_data |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")), 
    ggplot2::aes(x = age, y = functional_maturity)) +
  ggplot2::geom_point() +
  ggplot2::ylab("Functional Maturity") +
  ggplot2::geom_vline(xintercept = 7, color = "blue", linetype = 2) + 
  ggplot2::xlab("Age years") +
  ggplot2::theme_bw() +
  ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")) +
  ggplot2::facet_wrap(Area ~ .)

```

\newpage

```{r}
#| label: fig-maturity-ogive
#| echo: false
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated maturity-at-age by area and the biomass weighted coastwide (Spatial) maturity curve."
#| fig-alt: "Maturity in the south is later compared to fish in the north with the biomass weighted maturity curve being slightly right shifted compared to the north maturity curve."

load(here::here("data", "spatial_maturity.rda"))

ggplot2::ggplot(spatial_maturity, ggplot2::aes(x = age, y = p, color = Model)) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::ylab("Probability Mature") +
  ggplot2::xlab("Age years") +
  ggplot2::scale_colour_viridis_d() + 
  ggplot2::theme_bw() +
  ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white"))

```

\newpage


## Model Bridging

## Biology


### Selectivity


## Fits to Data


## Time-series


```{r} 
#| label: fig-sb
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of spawning biomass for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts7_Spawning_output_with_95_intervals.png")
)

```


```{r} 
#| label: fig-depl
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fraction of unfished spawning biomass for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts9_Relative_spawning_biomass_intervals.png")
)

```


```{r} 
#| label: fig-recruits
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of age-0 recruits for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png")
)

```

\pagebreak


```{r} 
#| label: fig-recdevs
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of recruitment deviations for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "recdevs2_withbars.png")
)

```



```{r} 
#| label: fig-biasramp
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Bias adjustment applied to the recruitment deviations (red line). Points are transformed variances relative to the assumed variance of recruitment."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "recruit_fit_bias_adjust.png")
)

```


```{r} 
#| label: fig-spr
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated time series of fishing intensity for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR3_ratiointerval.png")
)

```


```{r} 
#| label: fig-spr-phase
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Phase plot of fishing intensity versus fraction unfished for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "SPR4_phase.png")
)

```


```{r} 
#| label: fig-yield
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Estimated yield curve with reference points for the base model."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "yield2_yield_curve_with_refpoints.png")
)

```



```{r} 
#| label: fig-dyn-b0
#| echo: false 
#| warning: false 
#| eval: true 
#| fig-cap: "Dynamic B0 plot. The lower line shows the time series of estimated spawning output in the presence of fishing mortality. The upper line shows the time series that could occur under the same dynamics (including deviations in recruitment), but without fishing. The point at the left represents the unfished equilibrium."
#| fig-alt: ""
#| fig-pos: H

knitr::include_graphics(
  file.path(model_dir, "plots", "ts_DynamicB0.png")
)

```


## Sensitivity Analyse and Retrospectives


## Likelihood Profiles


## Reference Points and Forecasts

