---
title: 'Status of the sablefish (*Anoplopoma fimbria*) off the U.S. West Coast in 2025'
author:
  - name: 'Chantel R. Wetzel'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112'
  - name: 'Aaron M. Berger'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2032 SE OSU Drive Building 955'
        city: 'Newport, OR'
        state: 'OR'
        postal-code: '97365'
  - name: 'Cheryl Barnes'
    affiliations: 
      - name: 'Oregon State University'
        address: '2030 SE Marine Science Drive'
        city: 'Newport, OR'
        state: 'OR'
        postal-code: '97365'
  - name: 'Joshua A. Zahner'
    affiliations:
      - name: 'University of Alaska Fairbanks'
        address: 'Department of Fisheries at Lena Point, College of Fisheries and Ocean Sciences, University of Alaska Fairbanks, 17101 Point Lena Loop Rd'
        city: 'Juneau, AK'
        state: 'AK'
        postal-code: '99801'
  - name: 'Nick Tolimieri'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112-2097'
  - name: 'Eric J. Ward'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112-2097'
  - name: 'Melissa Head'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98112-2097'
date: today
lang: en 
keep-tex: true 
cover: 
pdf-engine: lualatex
format: 
  pdf: 
    documentclass: scrartcl
    number-sections: true
    template-partials:
      - 'support_files/before-body.tex'
      - 'support_files/_titlepage.tex'
    include-in-header:
      - 'support_files/in-header.tex'
      - 'support_files/figures.tex'
      - 'support_files/pfmc.tex'
    toc: true
    lof: false
    lot: false
    titlepage-geometry: 
      - top=2in
      - bottom=1in
      - right=1in
      - left=1in
    link-citations: true
    fig-pos: 'H'
output-file: 'sablefish_SAR_2025' 
params:
   office: 'NWFSC'
   species: 'sablefish'
   spp_latin: 'Anoplopoma fimbria'
   region: 'U.S. West Coast'
bibliography: 
-  asar_references.bib 
---

\pagenumbering{gobble}

```{r, results='asis'}
#| label: 'load_packages'
#| eval: true
#| echo: false
#| warning: false
#| message: false
a <- knitr::knit_child('001_load_packages.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'load_functions'
#| eval: true
#| echo: false
#| warning: false
#| message: false

#' Create r4ss plots required for the report document
#'
#' @param  dir Directory to save figures. The full directory is required and if 
#'   provided a folder will be created in this location where figures will be 
#'   saved.  
#' @param  model_name String model folder name.
#' @param  fleetnames List of user specific fleet names to use in the figures or 
#'   tables.
#'
#' @author Chantel Wetzel
#' @export
#'
create_model_plots <- function(
  model_name,
  dir = here::here("model")
  ) {

  model_dir <- file.path(dir, model_name)
  model_output <- r4ss::SS_output(model_dir, verbose = FALSE, printstats = FALSE, hidewarn = TRUE)
  save(model_output, file = file.path(model_dir, "model_output.rda"))
  r4ss::SS_plots(
    replist = model_output
  )
}
#'
#' Create r4ss rda files required for the report document
#'
#' @param  dir Directory to save figures. The full directory is required and if 
#'   provided a folder will be created in this location where figures will be 
#'   saved.  
#' @param  model_name String model folder name.
#' @param  fleetnames List of user specific fleet names to use in the figures or 
#'   tables.
#'
#' @author Chantel Wetzel
#' @export
#'
create_model_tables <- function(
    model_name,
    dir = here::here("model"),
    fleetnames = c(
      "Trawl", 
      "Hook-and-Line", 
      "Pot",
      "Trawl Discards", 
      "Hook-and-Line Discards",
      "Pot Discard",
      "Triennial Early", 
      "Triennial Late", 
      "NWFSC Slope", 
      "WCGBT",
      "Recruitment Index"
    )
) {
  
  model_dir <- file.path(dir, model_name)
  model_output <- r4ss::SS_output(model_dir)
  r4ss::table_all(
    replist = model_output,
    fleetnames = fleetnames
  )
  r4ss::table_config(replist = model_output)
}

```


```{r} 
#| label: 'output_and_quantities'
#| echo: false 
#| warning: false 
#| eval: true 
#| message: false

spp <- "sablefish"
Spp <- "Sablefish"

# Define the location and name of the base model
model_name <- "8.36_base_model"
general_dir <- here::here("model", "base_model")
model_dir <- file.path(general_dir, model_name)

# pak::pak("r4ss/r4ss@wtatage_figs")
# Create revised weight-at-age figures
if (any(c(
  !file.exists(file.path(model_dir, "model_output.rda")), 
  !file.exists(file.path(model_dir, "plots"))
  ))) {
  create_model_plots(
    model_name = model_name,
    dir = general_dir
  )
  r4ss::SS_plots(
    replist = model_output, 
    plot = 1, 
    minyr = 2000, 
    maxyr = 2025,
    pheight_tall = 3.5, 
    pwidth = 6.5
  )
  # Move select r4ss figures
  invisible(file.copy(file.path(model_dir, "plots", "SPR4_phase.png"), here::here("report", "figures", "SPR4_phase.png"), overwrite = TRUE))
  invisible(file.copy(file.path(model_dir, "plots", "ts7_Spawning_output_with_95_intervals.png"), here::here("report", "figures",     "ts7_Spawning_output_with_95_intervals.png"), overwrite = TRUE) )
  invisible(file.copy(file.path(model_dir, "plots", "yield2_yield_curve_with_refpoints.png"), here::here("report", "figures",   "yield2_yield_curve_with_refpoints.png"), overwrite = TRUE))
}
load(file.path(model_dir, "model_output.rda"))

table_dir <- file.path(model_dir, "tables")
if (!file.exists(table_dir)) {
  create_model_tables(
    model_name = model_name,
    dir = general_dir
  )
}

# Other potential directories that could be useful
sens_dir <-  here::here("model", "_sensitivities")
retro_dir <- profile_dir <- general_dir

# Define objects from the list created by r4ss::SS_output
end_year <- model_output$endyr + 1
data_end_year <- model_output$endyr
start_year <- model_output$startyr
all_years <- start_year:end_year
proj_years <- (end_year + 2):(end_year + 12)
es_10_years <- (end_year - 10):data_end_year
es_12_years <- (end_year - 10):(end_year + 1)

# Define some derived quantities
sb <- model_output$derived_quants[model_output$derived_quants$Label %in% paste0("SSB_", all_years), "Value"]
sb0 <- model_output$derived_quants[model_output$derived_quants$Label == "SSB_Virgin", "Value"] 
depl <- round(sb / sb0, 3)
quants <- data.frame(
  year = all_years, 
  sb = round(sb, 0), 
  depl = depl)

depl_final <- quants[quants$year == end_year, 'depl']
quants[, "sb"] <- format(round(quants[, "sb"], 0), scientific = FALSE, big.mark = ",")
sb_final <- quants[quants$year == end_year, 'sb'] 
min_sb <- min(quants[, "sb"])
min_sb_year <- quants[quants$sb == min_sb, 'year']

min_depl <- min(quants$depl)
min_depl_year <- quants[quants$depl == min_depl, 'year']

catch_spr <- model_output$derived_quants |>
  dplyr::filter(Label == "Dead_Catch_SPR") |>
  dplyr::mutate(
    catch_spr = round(Value, 0)
  ) |>
  dplyr::select(catch_spr) |>
  as.numeric() |>
  format(scientific = FALSE, big.mark = ",")

sb_spr <- model_output$derived_quants |>
  dplyr::filter(Label == "SSB_Btgt") |>
  dplyr::mutate(
    sb_spr = round(Value, 0)
  ) |>
  dplyr::select(sb_spr) |>
  as.numeric() |>
  format(scientific = FALSE, big.mark = ",")

# Calculated harvest metrics
fixed_catch <- round(model_output$derived_quants[model_output$derived_quants$Label %in% paste0("ForeCatch_", end_year:(end_year + 1)), "Value"], 0) |>
  format(scientific = FALSE, big.mark = ",")
est_ofl <- round(model_output$derived_quants[model_output$derived_quants$Label %in% paste0("OFLCatch_", proj_years), "Value"], 0) 
est_abc <- round(model_output$derived_quants[model_output$derived_quants$Label %in% paste0("ForeCatch_", proj_years), "Value"], 0) 

# Parameter values
recdevs <- dplyr::bind_rows(
  model_output$parameters |>
    dplyr::filter(grepl("Main_RecrDev", Label)) |>
    dplyr::select(Value),
  model_output$parameters |>
    dplyr::filter(grepl("Late_RecrDev", Label)) |>
    dplyr::select(Value)
  )
rec_years <- readr::parse_number(rownames(recdevs))
recdevs <- data.frame(
  years = rec_years,
  recdevs = recdevs$Value
)
high_5 <- sort(recdevs[, "recdevs"], decreasing = TRUE, index.return = TRUE)$ix[1:5]
top_recdevs <- recdevs[high_5, ]
low_5 <- sort(recdevs[, "recdevs"], decreasing = FALSE, index.return = TRUE)$ix[1:5]
low_recdevs <- recdevs[low_5, ]

sum_main_devs <- model_output$parameters |>
  dplyr::filter(grepl("Main_RecrDev", Label)) |>
  dplyr::select(Value) |>
  dplyr::summarise(sum = round(sum(Value), 2)) |>
  as.numeric()

# Parameters: biology values
m <- model_output$parameters |>
  dplyr::filter(Label %in% c("NatM_uniform_Fem_GP_1", "NatM_uniform_Mal_GP_1")) |>
  dplyr::select(Label, Value) |>
  dplyr::mutate(Value = round(Value, 3)) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric()
h <- model_output$parameters |>
  dplyr::filter(Label %in% c("SR_BH_steep")) |>
  dplyr::select(Value) |>
  as.numeric()
Linf <- model_output$parameters |>
  dplyr::filter(Label %in% c("L_at_Amax_Fem_GP_1", "L_at_Amax_Mal_GP_1")) |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(1)
Lmin <- model_output$parameters |>
  dplyr::filter(Label %in% c("L_at_Amin_Fem_GP_1", "L_at_Amin_Mal_GP_1")) |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(1)
k <- model_output$parameters |>
  dplyr::filter(Label %in% c("VonBert_K_Fem_GP_1", "VonBert_K_Mal_GP_1")) |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(3)


sigma_r <- model_output$parameters |>
  dplyr::filter(Label == "SR_sigmaR") |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(2)

find <- grep("Age_DblN_peak_TWL_Discards", model_output$parameters$Label)
trawl_discard_peak <- model_output$parameters[find, ] |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(2)

find <- grep("Age_DblN_peak_HKL_Discards", model_output$parameters$Label)
hkl_discard_peak <- model_output$parameters[find, ] |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(2)

find <- grep("Age_DblN_peak_Pot_Discard", model_output$parameters$Label)
pot_discard_peak <- model_output$parameters[find, ] |>
  dplyr::select(Label, Value) |>
  tidyr::pivot_wider(
    names_from = Label,
    values_from = Value
  ) |>
  as.numeric() |>
  round(2)

# Survey catchabilities and added sd
find <- grep("Q_base", model_output$parameters$Label)
q <- model_output$parameters[find, ] |>
  dplyr::select(Label, Value) 
rownames(q) <- NULL
q[, "Value"] <- c(round(exp(q[1:4, "Value"]), 2), round(q[5, "Value"], 2))
find <- grep("Q_extraSD", model_output$parameters$Label)
survey_add_sd <- model_output$parameters[find, ] |>
  dplyr::select(Label, Value) |>
  dplyr::mutate(Value = round(Value, 2))
rownames(survey_add_sd) <- NULL

ofl_sigma <- round(model_output$OFL_sigma, 2)
sb_sigma <- round(model_output$Pstar_sigma,2)

# Read in adopted harvest specifications:
cw_harvest_spex <- read.csv(here::here("data-processed", "harvest_spex.csv")) |>
  dplyr::filter(AREA_NAME == "CW", YEAR %in% es_12_years) |>
  dplyr::mutate(
    VAL = round(VAL, 0)
  )
area_harvest_spex <- read.csv(here::here("data-processed", "harvest_spex.csv")) |>
  dplyr::filter(AREA_NAME != "CW", YEAR %in% es_12_years) |>
  dplyr::mutate(
    VAL = round(VAL, 0)
  )

# read in maturity data
load(here::here("data", "maturity_data.rda"))
load(here::here("data", "maturity.rda"))
load(here::here("data", "spatial_maturity.rda"))
maturity_data <- maturity_data  |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")
  )
# load the previous model output
load(here::here("data", "model_output_2023.rda"))

ctl_file <- r4ss::SS_readctl(file = file.path(model_dir, "control.ss_new"))
gmt_catch <- r4ss::SS_ForeCatch(model_output) |>
  dplyr::filter(`#Year` %in% c(2025, 2026)) |>
  dplyr::rename(Year = `#Year`)

``` 

{{< pagebreak >}} 

Please cite this publication as: 

Wetzel, C.R., A.M. Berger, C. Barnes, J.A. Zahner, N. Tolimieri, E.J. Ward., and M. Head. 2025. Status of sablefish (*Anoplopoma fimbria*) off the U.S. West Coast in 2025. Pacific Fisheries Management Council, Portland, Oregon.

 {{< pagebreak >}} 
 
\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}


```{r, results='asis'}
#| label: 'load_tables'
#| eval: true
#| echo: false
#| warning: false
#| message: false
a <- knitr::knit_child('002_load_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'disclaimer'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('00a_disclaimer.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

\printnoidxglossaries

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'one_page'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('01_one_page.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'executive_summary'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('01_executive_summary.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

{{< pagebreak >}} 

\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\setcounter{section}{0}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

```{r, results='asis'}
#| label: 'introduction'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('02_introduction.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('03_data.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'assessment-configuration'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04a_assessment-configuration.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'assessment-results'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04b_assessment-results.qmd', quiet = TRUE)
cat(a, sep = '\n')
```


```{r, results='asis'}
#| label: 'assessment-sensitivity'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04c_assessment-sensitivity.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

```{r, results='asis'}
#| label: 'assessment-historical'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04d_historical.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 
 

```{r, results='asis'}
#| label: 'management'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('05_management.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'acknowledgments'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('06_acknowledgments.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'references'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('07_references.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('08_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'figures'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('09_figures.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'appendix'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('11_appendix.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

