
## Stock Structure

The 2025 stock assessment will report the status of sablefish off the U.S. West Coast as a single population unit. In this regard, there are several sablefish management and population dynamic features that the STAT team has considered, and some that the team will continue to consider during the development of assessment models. Genetically speaking, the U.S. West Coast area for sablefish can be considered a sub-stock given the genetic similarity between populations spanning from Southern California to Western Alaska [(Jasonowicz et al., 2017)](https://cdnsciencepub.com/doi/full/10.1139/cjfas-2016-0012). The STAT team will continue to utilize the Pacific Fishery Management Council default coast-wide (U.S. West Coast) stock delineation for purposes of the 2025 stock assessment.

Analyses investigating heterogeneity in sablefish biology and life history, both within the U.S. West Coast area and among neighboring Northeast Pacific sub-stock areas (British Columbia and Alaska), have indicated important population dynamic gradients across these areas. Patterns in sablefish growth are evident across their range, with southern areas generally having smaller maximum body size and intrinsic growth rates. Along the U.S. West Coast, a significant break-point in growth was detected at 36&deg;N latitude [(Kapur et al., 2020)](https://www.sciencedirect.com/science/article/pii/S0165783619302693).

```{r}
#| label: kapur_growth
#| echo: false
#| fig-cap: "Distribution of sablefish and major breakpoints in growth (areas demarcated by red dashed lines) in the Northeast Pacific (reproduced from Kapur et al., 2020; Fish.Res. 105414)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "kapur_growth.png")
)

```

Recruitment is widespread along the shelf-slope break. A single stock-recruitment function will be used to characterize sablefish productivity specific to the U.S. West Coast area. However, the rate at which females mature and become reproductively active can differ with latitude. Please refer to the [Biology](https://connect.fisheries.noaa.gov/us_west_coast_sablefish_2025/04_biology.html) section for further details on both growth and maturity.

Sablefish are well known to move considerable distances. Recent transboundary (Alaska, British Columbia, and U.S. West Coast) analyses of sablefish population dynamics indicated the presence of non-zero movement rates across different areal and stock definitions [(Kapur et al., 2024)](https://cdnsciencepub.com/doi/10.1139/cjfas-2024-0008). While movement estimates between the U.S. West Coast and British Columbia were similar (see the red circles), there remains a net difference due to immigration and emigration because of the difference in area-specific population size for which the rates apply.

```{r}
#| label: kapur_transb
#| echo: false
#| fig-cap: "Movement rate estimates for different area and stock definitions as part of simulation analyses looking into the influence of range-wide management outcomes (reproduced from Kapur et al., 2024; CJFAS 101139)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "kapur_transb.png")
)

```

At this time, the STAT team anticipates ignoring movement to (and from) British Columbia given the relatively low rates. The team may explore sensitivity models that increase natural mortality to account for estimates of net fish loss due to emigration. The work of Kapur et al. (2024) highlighted that the biggest potential risk for sablefish throughout their Eastern Pacific range was associated with localized depletion in isolated stock components in the southern U.S. West Coast area. 

## Regional Trends -- Including Canada and Alaska

Exploring regional trends in population dynamics can be a useful way to elicit patterns and gain insights, particularly for stocks such as sablefish that are connected over long distances spanning multiple management jurisdictions. Landings data, population dynamic trends, and stock status are compared between the most recent sablefish stock assessments for the U.S. West Coast [(Johnson et al., 2023)](https://www.pcouncil.org/documents/2024/02/status-of-sablefish-anoplopoma-fimbria-along-the-u-s-west-coast-in-2023.pdf/), British Columbia [(DFO, 2023)](https://waves-vagues.dfo-mpo.gc.ca/library-bibliotheque/41102587.pdf), and Alaska [(Goethel and Cheng, 2024)](https://www.npfmc.org/wp-content/PDFdocuments/SAFE/2024/Sablefish.pdf).

Landings of sablefish between stock assessment areas has changed considerably over time, but generally all three areas saw a decline in catch during the 1990s to 2000s. There has been a slight uptick in catch in recent years, especially in Alaska and British Columbia. For comparison, catch in 2021 was nearly 22 mt in Alaska, 2.4 mt in British Columbia, and 5 mt along the U.S. West Coast.

```{r}
#| label: reg_landing
#| echo: false
#| fig-cap: "Landings of sablefish in Alaska, British Columbia, and along the U.S. West Coast from the most recent stock assessments. (Alaska and British Columbia data courtesy of Goethel and Cheng (2024) and DFO (2023), respectively)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "reg_landing.png")
)

```

Spawning biomass in Alaska and British Columbia have generally declined until the late 2010s, but are now increasing quite rapidly. The U.S. West Coast population generally declined from the 1970s to the early 2000s. The population was then relatively stable through the late 2010s, before increasing rapidly in recent years similar to Alaska and British Columbia. The current stock status (depletion) estimates are 67% of equilibrium biomass in Alaska in 2024, 54% in British Columbia (2022), and 63% for the U.S. West Coast area (2023).

```{r}
#| label: reg_ssb
#| echo: false
#| fig-cap: "Time series of sablefish spawning biomass estimates (with uncertainty intervals) in Alaska, British Columbia, and along the U.S. West Coast from the most recent stock assessments. (Alaska and British Columbia data courtesy of Goethel and Cheng (2024) and DFO (2023), respectively)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "reg_ssb.png")
)

```

In general, there are periods over time that coincide with above average recruitment across all three assessment areas. In particular, there has been several strong recruitment events in all three areas over the last 10 years. The three largest recruitment events (all scaled to age-0) over recent years were:

-   Alaska: 2016, 2017, 2019
-   British Columbia: 2015, 2016, 2017
-   U.S. West Coast: 2016, 2020, 2021

```{r}
#| label: reg_rec
#| echo: false
#| fig-cap: "Time series of sablefish recruitment estimates (with uncertainty intervals) in Alaska, British Columbia, and along the U.S. West Coast from the most recent stock assessments. (Alaska and British Columbia data courtesy of Goethel and Cheng (2024) and DFO (2023), respectively)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "reg_rec.png")
)

```

Overall stock status (at the time of the last assessment) in all three regions can be generalized to indicate that overfishing is not occurring and the stocks are not overfished. Each region has been above fishing-based reference levels and below biomass-based reference levels at one point or another during the time series. However, recent strong recruitment events are now contributing to spawning biomass, which is driving the trend of increasing stock status measures over recent years in all three areas.


```{r}
#| label: reg_status
#| echo: false
#| fig-cap: "Time series of sablefish biomass-based and fishing-based estimates of status in Alaska, British Columbia, and along the U.S. West Coast from the most recent stock assessments. (Alaska and British Columbia data courtesy of Goethel and Cheng (2024) and DFO (2023), respectively)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "reg_status.png")
)

```


**Disclaimer: All data summaries and exploration presented here are preliminary and may not be indicative of the final data that will be incorporated in the 2025 assessment**


