
## Stock Structure

The 2025 stock assessment will report the status of sablefish off the U.S. West Coast as a single population unit. In this regard, there are several sablefish management and population dynamic features that the STAT team has considered, and some that the team will continue to consider during the development of assessment models. Genetically speaking, the U.S. West Coast area for sablefish can be considered a sub-stock given the genetic similarity between populations spanning from Southern California to Western Alaska (Jasonowicz et al. 2017). The STAT team will continue to utilize the Pacific Fishery Management Council default coast-wide (U.S. West Coast) stock delineation for purposes of the 2025 stock assessment.

Analyses investigating heterogeneity in sablefish biology and life history, both within the U.S. West Coast area and among neighboring Northeast Pacific sub-stock areas (British Columbia and Alaska), have indicated important population dynamic gradients across these areas. Patterns in sablefish growth are evident across their range, with southern areas generally having smaller maximum body size and intrinsic growth rates. Along the U.S. West Coast, a significant break-point in growth was detected at 36&deg;N latitude (Kapur et al. 2020).

```{r}
#| label: kapur_growth
#| echo: false
#| fig-cap: "Distribution of sablefish and major breakpoints in growth (areas demarcated by red dashed lines) in the Northeast Pacific (reproduced from Kapur et al., 2020; Fish.Res. 105414)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "kapur_growth.png")
)

```

Recruitment is widespread along the shelf-slope break. A single stock-recruitment function will be used to characterize sablefish productivity specific to the U.S. West Coast area. However, the rate at which females mature and become reproductively active can differ with latitude. Please refer to the Biology section for further details on both growth and maturity.

Sablefish are well known to move considerable distances. Recent transboundary (Alaska, British Columbia, and U.S. West Coast) analyses of sablefish population dynamics indicated the presence of non-zero movement rates across different areal and stock definitions (Kapur et al., 2024). While movement estimates between the U.S. West Coast and British Columbia were similar (see the red circles), there remains a net difference due to immigration and emigration because of the difference in area-specific population size for which the rates apply.

```{r}
#| label: kapur_transb
#| echo: false
#| fig-cap: "Movement rate estimates for different area and stock definitions as part of simulation analyses looking into the influence of range-wide management outcomes (reproduced from Kapur et al., 2024; CJFAS 101139)."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "kapur_transb.png")
)

```

At this time, the STAT team anticipates ignoring movement to (and from) British Columbia given the relatively low rates. The team may explore sensitivity models that increase natural mortality to account for estimates of net fish loss due to emigration. The work of Kapur et al. (2024) highlighted that the biggest potential risk for sablefish throughout their Eastern Pacific range was associated with localized depletion in isolated stock components in the southern U.S. West Coast area. 

## Trends Observed in Canada and Alaska


**Disclaimer: All data summaries and exploration presented here are preliminary and may not be indicative of the final data that will be incorporated in the 2025 assessment**


