```{r}
#| label: load-fishery-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
# project_dir <- here::here()
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")
load(file.path(data, "data_commercial_catch.rda"))

landings <- data_commercial_catch |>
  dplyr::mutate(
    gear_group = dplyr::case_when(
      gear_group %in% c("hkl") ~ "HKL",
      gear_group %in% c("pot") ~ "Pot",
      TRUE ~ "Trawl")
  ) |>
  dplyr::rename(Gear = gear_group)

landings_by_state <- data_commercial_catch |>
  dplyr::filter(year >= 2011, state != "at-sea") |>
  dplyr::mutate(total = sum(catch_mt)) |>
  dplyr::group_by(state) |>
  dplyr::summarise(
    all_states = unique(total),
    by_state = sum(catch_mt),
    prop = 100 * round(by_state / all_states, 2)
  ) 

landings_by_gear <- data_commercial_catch |>
  dplyr::filter(year >= 2011, state != "at-sea") |>
  dplyr::mutate(total = sum(catch_mt)) |>
  dplyr::group_by(gear_group) |>
  dplyr::summarise(
    all_gear = unique(total),
    by_gear = sum(catch_mt),
    prop = 100 * round(by_gear / all_gear, 2))

discard_rates <- read.csv(
  file.path(data_proc, "data_commercial_discard_rates.csv")) |>
  dplyr::mutate(
    Fleet = dplyr::case_match(fleet, 2 ~ "HKL",
                              3 ~ "Pot",
                              1 ~ "Trawl"),
    lwr = qnorm(0.025, discard_rate, sd),
    upr = qnorm(0.975, discard_rate, sd)
  ) |>
  dplyr::mutate(
    lwr = dplyr::case_when(lwr < 0 ~ 0, .default = lwr),
    year_offset = dplyr::case_when(fleet == 1 ~ year, fleet == 2 ~ year + 0.25, .default = year + 0.5)
  ) 

```

## Model and Fleet Structure

For the 2025 assessment, we plan to have three coastwide fleets separated by the following gear types: trawl, hook-and-line (HKL), and pot.  The HKL and pot gears will be broken into two separate fleets for initial modeling, to ensure that we can capture different selectivities and retention practices across time, if appropriate. This proposed fleet structure varies from the fleet structure used in the most recent assessments of sablefish that combined HKL and pot gear into a single fixed-gear fleet.

## Landings Data

The historical landings (pre-1981) for sablefish remain similar to the historical landings used in the most recent assessment (e.g., 2023). Landings for sablefish are based on data from state-specific historical reconstructions, historical foreign fleets (1966-1980), PacFIN, at-sea bycatch estimates (1978-2024), and recreational fishing (2001-2024). 

```{r}
#| label: landings-recent
#| echo: false
#| warning: false
#| fig-cap: "Coastwide landings (mt) of sablefish by gear since 1981. Note, the 2024 landings are incomplete."
#| fig-cap-location: margin

ggplot(landings |> dplyr::filter(year >= 1981), aes(x = year, y = catch_mt, fill = Gear)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("Year") + ylab("Landings (mt)") +
  scale_y_continuous(
    labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  scale_fill_viridis_d(begin = 0.0, end = 0.5)

```

Sablefish are caught coastwide. Since 2011, the total percentage of sablefish being landing with state ports are `r landings_by_state[1, "prop"]`% in California, `r landings_by_state[2, "prop"]`% in Oregon, and `r landings_by_state[3, "prop"]`% in Washington.  Additionally, since 2011, the total percentage of sablefish being landing by gear are `r landings_by_gear[1, "prop"]`% by HKL, `r landings_by_gear[2, "prop"]`% by pot, and `r landings_by_gear[3, "prop"]`% by trawl.

```{r}
#| label: landings-state
#| echo: false
#| warning: false
#| fig-cap: "Sablefish landings (mt) by state. Note, this does not include that at-sea bycatch and the 2024 landings are incomplete."
#| fig-cap-location: margin

ggplot(landings |> 
         filter(!state %in% c("at-sea"), year > 1980), 
       aes(x = year, y = catch_mt, fill = state)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("Year") + ylab("Landings (mt)") +
  scale_y_continuous(
    labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  scale_fill_viridis_d(begin = 0.5, end = 1)

```

Across the last ten years, approximately 41%, 31%, and 28% of sablefish landings have been by HKL, trawl, and pot gear, respectively.

```{r}
#| label: landings-table
#| warning: false
#| echo: false
#| tbl-cap: "Coastwide landings (mt) by gear and total landings for the last ten years.  Note, the landings for 2024 are incomplete."
#| tbl-cap-location: margin

landings |>
  dplyr::filter(year >= 2015) |>
  dplyr::group_by(year, Gear) |>
  dplyr::summarise(
    mt = sum(catch_mt),
  ) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    total = sum(mt)
  ) |>
  dplyr::select(year, Gear, mt, total) |>
  tidyr::pivot_wider(
    names_from = Gear,
    values_from = mt
  ) |>
  dplyr::relocate(total, .after = Trawl) |>
  dplyr::rename(
    Year = year,
    `HKL (mt)` = HKL,
    `Pot (mt)` = Pot,
    `Trawl (mt)` = Trawl,
    `Total Landings (mt)` = total
  ) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Coastwide landings (mt) by gear and total landings for the last ten years.  Note, the landings for 2024 are incomplete."
  ) |>
  fmt_number(
    columns = 2:5,
    decimals = 0
  ) |>
  cols_align(
    align = "center"
  ) |>
  #cols_width(
  #  everything() ~ px(10)
  #) |>
  tab_options(
    table.font.size = 20,
    table.align = "right"
  ) |>
  tab_style(
    style = list(
      cell_text(style = "italic")
    ),
    locations = cells_body(
      rows = Year > 2023
    )
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

## Discard Data

Similar to previous assessments for sablefish, we plan on modeling selectivity (all fish caught by gear), retention (fish landed), and mortality for discarded fish. The model will assume that 50% discard mortality for fish caught by trawl gear and a 20% discard mortality for HKL and pot gear. These discard mortality rates were included in the Groundfish Expanded Multi-year Mortality (GEMM) estimates produced by the Fisheries Observation Science (FOS) starting in 2019. Discarding is informed by data collected by the West Coast Groundfish Observer Program (WCGOP). Data are available starting from 2002 to 2023. WCGOP collects information on the size and amount of fish discarded. Starting in 2011, vessels in the catch share program have 100% observer coverage (or are observed via electronic monitoring). For non-catch share vessels, the proportion of total trips observed varies by sector.

Average discarding rate by gear type are:

-   HKL gear range between `r round(min(discard_rates[which(discard_rates$Fleet == "HKL" & discard_rates$year < 2011), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "HKL" & discard_rates$year < 2011), "discard_rate"]), 2)` from 2002-2010 and `r round(min(discard_rates[which(discard_rates$Fleet == "HKL" & discard_rates$year >= 2011), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "HKL" & discard_rates$year >= 2011), "discard_rate"]), 2)` since 2011.

-   Pot gear range between `r round(min(discard_rates[which(discard_rates$Fleet == "Pot" & discard_rates$year < 2011), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "Pot" & discard_rates$year < 2011), "discard_rate"]), 2)` from 2002-2010 and `r round(min(discard_rates[which(discard_rates$Fleet == "Pot" & discard_rates$year >= 2011), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "Pot" & discard_rates$year >= 2011), "discard_rate"]), 2)` since 2011.

-   Trawl gear range between `r round(min(discard_rates[which(discard_rates$Fleet == "Trawl" & discard_rates$year < 2011 & discard_rates$year >= 2002), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "Trawl" & discard_rates$year < 2011 & discard_rates$year >= 2002), "discard_rate"]), 2)` from 2002-2010 and `r round(min(discard_rates[which(discard_rates$Fleet == "Trawl" & discard_rates$year >= 2011), "discard_rate"]), 2)` to `r round(max(discard_rates[which(discard_rates$Fleet == "Trawl" & discard_rates$year >= 2011), "discard_rate"]), 2)` since 2011. Observed discarding has been increasing annually since 2019.

```{r}
#| label: discard-rates
#| echo: false
#| warning: false
#| fig-cap: "Coastwide discard rates by gear type based on West Coast Groundfish Observer data. For 2011 - 2023, discard rates were weighted by expanded yearly discard totals for catch share and non-catch share sectors using the Groundfish Expanded Multi-species Mortality report."
#| fig-cap-location: margin

ggplot(discard_rates |> dplyr::filter(year > 2000), aes(x = year, y = discard_rate)) +
  geom_point() +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = lwr, ymax = upr)) +
  theme_bw() +
  scale_color_viridis_d() +
  ylab("Discard Rates") + xlab("Year") +
  facet_grid("Fleet")

```

An additional discard study (referred to as the Pikitch study) was conducted in 1985-1987 on-board 40 groundfish trawl vessels operating out of Newport, Astoria, and Coos Bay, Oregon. The estimated discard rates based on this study are shown in the table below.

```{r}
#| label: pikitch-discard-rates
#| warning: false
#| echo: false
#| tbl-cap: "Estimated trawl discard rates based on the Pikitch discard study."
#| tbl-cap-location: margin

discard_rates |>
  dplyr::filter(year < 2000) |>
  dplyr::select(year, discard_rate, sd) |>
  dplyr::mutate(
    Gear = "Trawl"
  ) |>
  dplyr::rename(
    Year = year,
    `Discard Rate` = discard_rate,
    SD = sd
  ) |>
  dplyr::relocate(
    Gear, .after = Year
  ) |>
  gt() |>
  tab_header(
    title = "Estimated trawl discard rates based on the Pikitch discard study."
  ) |>
  fmt_number(
    columns = 3:4,
    decimals = 2
  ) |>
  cols_align(
    align = "center"
  ) |>
  #cols_width(
  #  everything() ~ pct(25)
  #) |>
  tab_options(
    table.font.size = 20,
    table.align = "center"#,
    #table.width = pct(50),
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

The Enhanced Data Collection Project (EDCP) conducted off Oregon for trawl vessels in 1997 collected discarding data from select vessels that estimated a discarding rate of 0.40. However, the discarding estimates from this study were found to not be representative of the fleet at large (Sampson, 2002). We currently do not plan to include the EDCP data due to concerns identified by Sampson (2002). Given that the EDCP discarding rate is similar to the rates found by the Pikitch study for trawl gear, excluding the EDCP data is likely to have little to no impact on the assessment.

Previous assessments have assumed there were changes in selectivity or retention for the following time periods:

-   Trawl gear:
    -   1942-1946: full retention of age-1+ sablefish during WWII;
    -   1982-2002: trip-limits induced discarding;
    -   2003-2010: spatial closures (Rockfish Conservation Areas);
    -   2011-2018: catch-share program; and
    -   2019+: discard mortality rates begin to be applied in the GEMM.
-   HKL and Pot gears:
    -   1942-1946: full retention of age-1+ sablefish during WWII;
    -   1996-2002: trip-limits induced discarding;
    -   2003-2010: spatial closures (Rockfish Conservation Areas);
    -   2011-2018: catch-share program; and
    -   2019+: discard mortality rates begin to be applied in the GEMM.

Questions:

-   What is behind the change in discarding observed for trawl gear starting in 2019?

-   Are the listed management changes listed above reasonable? Are there other changes in the fishery that should be considered?

## Fishery Compositions

Below are summaries of the available commercial length and age samples and comparisons of the length and age distribution by year, gear, and area in Pacific Fisheries Information Network (PacFIN).  It is important to note the Cooperative Ageing Lab are still hard at work reading otoliths from the commercial fishery and the age data are incomplete as of March 5th, 2025. The ageing lab is planning to fishing ageing by the data deadline of April 21st, 2025.

```{r}
#| label: load-commercial-bds-data
#| echo: false
#| warning: false
load(file.path(data, "data_commercial_bds.rda"))
bds_data <- data_commercial_bds |> 
  as_tibble() |>
  mutate(
    area = case_when(
      state %in% c("WA", "OR") ~ state,
      TRUE ~ paste(area, state)
    ),
    period = case_when(
      year > 2018 ~ "2019-2024",
      year < 2011 ~ "1980-2010",
      TRUE ~ "2011-2018"
    ),
    area = factor(area, levels=c("South CA", "North CA", "OR", "WA"))
  )
  
```

Length samples are generally available for all gear types across the three states with limited samples beginning in 1967 in Washington, 1978 in California, and 1987 in Oregon. 

```{r}
#| label: nsamples-length
#| echo: false
#| warning: false
#| fig-cap: "Number of available length samples over time, grouped by gear type (TWL = trawl, HKL = hook-and-line, POT = pot) and state."
#| fig-cap-location: margin

bds_data |> 
  filter(geargroup %in% c("HKL", "POT", "TWL"), !is.na(length_cm)) |>
  count(year, area, geargroup) |>
  ggplot(aes(x = year, y = n, fill = geargroup)) +
    geom_bar(stat = 'identity') +
    facet_wrap(facets = c("area")) +
    scale_fill_viridis_d() +
    labs(y="Number of Length Samples", x="Year", fill="Gear Type", main="") +
    theme_bw()
```

Age samples are generally available for all gear types across the three states, except for pot and hook-and-line gear in southern California.  Age data begin in 1986 for California and Washington and 1987 in Oregon.

```{r}
#| label: nsamples-age
#| echo: false
#| warning: false
#| fig-cap: "Number of available ages over time, grouped by gear type (TWL = trawl, HKL = hook-and-line, POT = pot) and state. Note, this does not include unread otoliths."
#| fig-cap-location: margin

bds_data |> 
  filter(geargroup %in% c("HKL", "POT", "TWL"), !is.na(age_years)) |>
  count(year, area, geargroup) |>
  ggplot(aes(x = year, y = n, fill = geargroup)) +
    geom_bar(stat = 'identity') +
    facet_wrap(facets = c("area")) +
    scale_fill_viridis_d() +
    labs(y="Number of Age Samples", x="Year", fill="Gear Type") +
    theme_bw()
```

Commercial length compositions from trawl fisheries are similar across the coast. Meanwhile, length compositions from the hook-and-line (HKL) and pot fisheries are skewed towards larger fish in Oregon and Washington as compared to California. Length compositions of hook-and-line and pot gear are often similar, though the pot fishery may a slightly reduced frequency of larger fish in Oregon.

```{r}
#| label: length-by-gear-region
#| echo: false
#| warning: false
#| fig-cap: "Distribution of length samples between 1967-2024 by gear type and state."
#| fig-cap-location: margin

bds_data |>
  group_by(area, geargroup) |>
  mutate(area = factor(area)) |>
  select(sex, length_cm, area, geargroup) |>
  ggplot(aes(x = length_cm, fill = area)) +
    ggridges::geom_density_ridges(aes(y=geargroup), alpha=0.5, scale=1) +
    labs(y = "Gear", x = "Length (cm)", fill = "Area")+
    scale_x_continuous(limits = c(18, 100), expand=c(0, 0)) +
    scale_fill_viridis_d() +
    theme_bw()
```

Commercial age compositions from trawl fisheries are generally similar across the coast, although there is a marginally lower frequency of older ages being selected in Washington. Similarly, there is a lower frequency of older ages being selected by pot gear in Washington compared to Oregon and Northern California. The distribution of ages caught by hook-and-line gear are generally similar across the coast. 

```{r}
#| label: age-by-gear-region
#| echo: false
#| warning: false
#| fig-cap: "Distribution of all age samples between 1986-2024 by gear type and state."
#| fig-cap-location: margin

bds_data |>
  group_by(area, geargroup) |>
  mutate(area = factor(area)) |>
  select(sex, age_years, area, geargroup) |>
  ggplot(aes(x=age_years, fill = area))+
    ggridges::geom_density_ridges(aes(y = geargroup), alpha=0.5, scale=1)+
    labs(y ="Gear", x="Age (years)", fill = "Area")+
    scale_x_continuous(limits = c(0, 50), expand = c(0, 0))+
    scale_fill_viridis_d()+
    theme_bw()
```

Length compositions within each state and by gear type have remained largely consistent over time (southern California for which there is a paucity available post-2011). 

```{r}
#| label: length-by-gear-region-period
#| echo: false
#| warning: false
#| fig-cap: "Distribution of length samples by gear and region for three time periods. Time periods are generally reflective of potential changes in retention: 2011 implementation of the catch share program and 2019 mortality credits begun being applied to discards in the catch share program."
#| fig-cap-location: margin

ggplot(bds_data,
       aes(x = length_cm, y = geargroup, fill = period, alpha = 0.25)) +
  ggridges::geom_density_ridges(scale = 0.9) +
  scale_x_continuous(limits = c(18, 100), expand=c(0, 0), breaks=seq(18, 100, 20)) +
  labs(y = "Gear", x = "Length (cm)", fill = "Time Period") +
  guides(alpha = "none") +
  facet_wrap(~area, ncol=2) +
  scale_fill_viridis_d() +
  theme_bw() +
  theme(
    panel.spacing.x = unit(30, "pt")
  )
```

```{r}
#| label: age-by-gear-region-period
#| echo: false
#| warning: false
#| fig-cap: "Distribution of age samples by gear and region for three time periods. Time periods are generally reflective of potential changes in retention: 2011 implementation of the catch share program and 2019 mortality credits begun being applied to discards in the catch share program."
#| fig-cap-location: margin

ggplot(bds_data,
       aes(x = age_years, y = geargroup, fill = period, alpha = 0.25)) +
  ggridges::geom_density_ridges(scale=0.9) +
  scale_x_continuous(limits=c(0, 50), expand=c(0, 0), breaks=seq(0, 50, 10))+
  labs(y = "Gear", x = "Age", fill = "Time Period")+
  guides(alpha="none")+
  facet_wrap(~area, ncol=2) +
  scale_fill_viridis_d()+
  theme_bw()+
  theme(
    panel.spacing.x = unit(30, "pt")
  )
```

**Disclaimer: All data summaries and exploration presented here are preliminary and may not be indicative of the final data that will be incorporated in the 2025 assessment**

