
```{r}
#| label: load-survey-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
library(nwfscSurvey)

#!!!!!!  Revise to your specific directory when trying to build !!!!!
# this needs to be fixed to use relative directories
project_dir <- "C:/Assessments/2025/sablefish_2025"

data_proc <- file.path(project_dir , "data-processed")
data <- file.path(project_dir , "data")

load(file.path(data, "data_survey_catch.rda"))
load(file.path(data, "data_survey_bio.rda"))
wcgbt_catch <- data_survey_catch$nwfsc_combo |>
  dplyr::mutate(
    positive = dplyr::case_when(total_catch_numbers > 0 ~ 1, .default = 0),
    state = dplyr::case_when(Latitude_dd > 46.25 ~ "WA", 
                             Latitude_dd < 36.0 ~ "SCA", 
                             Latitude_dd >= 36.0 & Latitude_dd < 42.0 ~ "NCA",
                             .default = "OR"))

wcgbt_bio <- data_survey_bio$nwfsc_combo

wcgbt_area <- read.csv(here::here("figures", "est_by_area.csv")) |>
  dplyr::filter(area %in%  c("North of Monterey Bay", "South of Monterey Bay")) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    total_est = sum(est)
  ) |>
  dplyr::mutate(
    prop = est / total_est
  )

wcgbt <- read.csv(here::here("figures", "est_by_area.csv")) |>
  dplyr::filter(area == "Coastwide") |>
  dplyr::mutate(
    ave_index = mean(est),
    index = est / ave_index,
    survey = "WCGBT"
  )

age_comps <- read.csv(file.path(data_proc, "data-survey-comps-ages-wcgbt.csv")) |>
  dplyr::select(-month, -fleet, -partition, -ageerr, -Lbin_lo, -Lbin_hi, -input_n) |>
  tidyr::pivot_longer(
    cols = 3:104,
    names_to = "age",
    values_to = "prop"
  ) |>
  dplyr::mutate(
    sex_chara = as.factor(substring(age, first = 1, last = 1)),
    sex_chara = dplyr::case_when(sex == 0 ~ "u", .default = sex_chara),
    sex_chara = as.factor(sex_chara),
    sex_type = dplyr::case_when(sex_chara == "u" ~ "u", .default = "b"),
    age_number = as.numeric(substring(age, first = 2, last = 3)),
    prop = round(prop / 100, 3)
  ) |>
  dplyr::select(-sex, -age) |>
  dplyr::rename(
    sex = sex_chara,
    age = age_number) 

```

## NWFSC West Coast Groundfish Bottom Trawl Survey

The NWFSC West Coast Groundfish Bottom Trawl (WCGBT) survey is the longest running annual trawl survey off the the U.S. West Coast.  The WCGBT survey is based on a random grid design; covering the coastal waters between 55 meter and 1,280 meter in depth sampling coastwide between U.S./Mexico and U.S./Canada border. Two vessels fish from north to south during each pass between late May and early October. The survey started in 2003 and has been conducted annually, except for 2020. Sablefish are one of the most commonly caught species in the WCGBT survey.

### Relative Indices of Abundance

A criticial data type produced by the WCGBT survey is a relative index of abundance that reflects potential changes across time in sablefish abundance. The index is estimated using a model that accounts for spatial and spatio-temporal effects. The index of abundance decreased between 2003-2009, then slight annual increases from 2010-2018, and then finally a steep increase from 2019-2022, and has remained relatively stable in the last few years of the survey.

```{r}
#| label: wcgbt-index
#| echo: false
#| fig-cap: "NWFSC West Coast Groundfish Bottom Trawl survey relative index of abundance (mt) for sablefish."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "index_coastwide.png")
)

```

The raw catch-per-unit-effort (CPUE) observed in the WCGBT survey data has a similar trend as the modeled index of abundance. The median CPUE by area reflects that the sharp increases in observations of sablefish in the survey is consistent in Washington, Oregon, and California north of 36 $^\circ$ N. latitude, but similar increases in CPUE has not been observed in southern California. 

```{r}
#| label: wcgbt-cpue
#| echo: false
#| fig-cap: "NWFSC West Coast Groundfish Bottom Trawl survey median catch per unit effort (CPUE) kilogram per kilometer squared (kg/km2) for sablefish."
#| fig-cap-location: margin
#| warning: false

cpue <- wcgbt_catch |>
  group_by(Year, state) |>
  summarise(
    med_cpue = median(cpue_kg_km2)
  )

ggplot(cpue, aes(x = Year, y = med_cpue, color = state)) +
  geom_line(linewidth = 1.5) +
  scale_color_viridis_d() +
  ylab("Median CPUE kg/km2") +
  theme_bw()

```

The model estimated biomass by areas north and south of 36 $^\circ$ N. latitude also reflects an increasing proportion of biomass in the northern area over the last 5 years.

```{r}
#| label: wcgbt-index-by-area
#| echo: false
#| tbl-cap: "Estimated percentage of sablefish biomass estimated north and south of 36 N. latitude based on the NWFSC WCGBT survey data over the last five years."
#| tbl-cap-location: margin

wcgbt_area |>
  filter(year >= 2019) |>
  mutate(
    Area = dplyr::case_when(area == "North of Monterey Bay" ~ "N. 36 Latitude",
                            .default = "S. 36 Latitude")
  ) |>
  rename(
    Year = year,
    Proportion = prop) |>
  select(Year, Area, Proportion) |>
  tidyr::pivot_wider(
    names_from = Area,
    values_from = Proportion
  ) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Estimated percentage of sablefish biomass estimated north and south of 36 N. latitude based on the NWFSC WCGBT survey data over the last five years."
  ) |>
  fmt_percent(
    columns = 2:3,
    decimals = 1
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20,
    table.align = "right"
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

The proportion of tows where sablfish are observed north of 36$^\circ$ N. latitude range between 60-80% and between 25-45% south of 36$^\circ$ N. latitude.

```{r}
#| label: presence-by-lat
#| echo: false
#| fig-cap: "Proportion of tows where sablefish was present or absent across all years by latitude for the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-cap-location: margin

nwfscSurvey::plot_proportion(
    data = wcgbt_catch |> dplyr::mutate(new = factor(cpue_kg_km2 <= 0, levels = c(FALSE, TRUE), labels = c("Present", "Absent"))),
    column_factor = new,
    column_bin = Latitude_dd,
    width = 1,
    boundary = 0,
    bar_width = "equal"
  )

```

The survey samples waters between 55-1,280 meters (30-700 fathoms). Sablefish are observed across all depths sampled with the shallowest depths having the lowest proportion of tows observing sablefish of 25-50% in waters shallower than 150 meters with the proportion of positive tows increasing with depth and peaking at greater than 95% positive between 950-1,050 meters.  

```{r}
#| label: presence-by-depth
#| echo: false
#| fig-cap: "Proportion of tows where sablefish was present or absent across all years by depth for the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-cap-location: margin

nwfscSurvey::plot_proportion(
    data = wcgbt_catch |>
      dplyr::mutate(
        new = factor(
          cpue_kg_km2 <= 0,
          levels = c(FALSE, TRUE),
          labels = c("Present", "Absent")
        )
      ),
    column_factor = new,
    column_bin = Depth_m,
    width = 50,
    boundary = 0,
    bar_width = "equal" 
  )
```


### Survey Compositions

By year the NWFSC West Coast Groundfish Bottom Trawl survey observes a slightly higher proportion of males compared to females.  

```{r}
#| label: prop-by-sex
#| echo: false
#| fig-cap: "Proportion of all NWFSC West Coast Groundfish Bottom Trawl survey tows that sampled sablefish between 2003 and 2024 (gold line). Sex-specific proportions of tows positive for sablefish are also shown for each year."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "proportion_by_sex.png")
)

```

Sablefish exhibit ontogenetic movement where fish move from shallow to deeper water as they grow and age. Across the coast, there is little evidence of segregation across all ages or sizes (e.g., young vs. old or small vs. large), with one notable exception in central California, around San Francisco, there observed population across depths is composed primarily of younger/smaller individuals across depths. However, there may be movement patterns of new cohorts that could be linked to areas of strong recruitment which are discussed further below.

```{r}
#| label: map-length-age
#| echo: false
#| fig-cap: "Sampling locations for sablefish collected by the NWFSC West Coast Groundfish Bottom Trawl survey between 2003 and 2024. Standard lengths (rounded down to the nearest cm) are shown on the left and age estimates (yr) are shown on the right. Lighter colors represent smaller or younger individuals and darker colors represent larger or older individuals. Latitudinal and inshore-offshore patterns were similar for males and females, thus sexes were combined."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "map-age-length.png")
)

```



```{r}
#| label: depth-length-age
#| echo: false
#| fig-cap: "Length (top; cm) and age (bottom; yr) compositions for sablefish collected by the NWFSC West Coast Groundfish Bottom Trawl Survey between 2003 and 2024 by sex and depth (m). Lighter colors represent smaller or younger individuals and darker colors represent larger or older individuals."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "length_age_compositions.png")
)

```


```{r}
#| label: map-length-age-percentile
#| echo: false
#| fig-cap: "Percentiles for sablefish collected by the NWFSC West Coast Groundfish Bottom Trawl survey between 2003 and 2024. Left: standard lengths (rounded down to the nearest cm). Right: age estimates (yr) are shown. Latitudinal patterns were similar for males and females, thus sexes were combined."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "Fig5.png")
)

```

Patterns of strong recruitments are visible in both the length and age composition data. For example, in 2008 there is mode of small fish observed around 25 cm that can be seen moving through the population in 2009 and 2010.  The strongest recruitment in recent years estimated in the 2023 assessment in descending order were 2020, 2021, 2016, 2008, and 2013.

```{r}
#| label: survey-length-comps
#| echo: false
#| fig-cap: "Proportion of lengths observed each year across sexes by the NWFSC West Coast Groundfish Bottom Trawl survey."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "length_proportion.png")
)

```

The proportion by age across years for estimated strong recruitment years can be variable from one year to the next. For example, the 2020 cohort that was first observed in 2021 was a large proportion of the observations of aged fish. In 2022, the proportion of observed age-2 fish decreased relative to the age-1 proportion in 2021, but the proportion of age-3 fish observed in 2023 increased relative to the proportion of age-2 fish in 2022. The variability in observations of year-classes across year can result in highly uncertain initial estimates of recruitment size that require additional years of data for the model to reliably estimate the year-class size.

```{r}
#| label: suvey-age-comp
#| echo: false
#| fig-cap: "Proportion of ages observed each year across sexes by the NWFSC West Coast Groundfish Bottom Trawl survey. Observed ages of 15 and higher were summed into a plus group of age 15."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "age_proportion.png")
)

```


### Other Trends in the Data

The [2024-25 California Current Ecosystem Status Report](https://www.pcouncil.org/documents/2025/02/f-1-a-cciea-report-1-2024-2025-california-current-ecosystem-status-report.pdf/) also included the estimated distribution of juvenile sablefish around select port groups in the NWFSC West Coast Groundfish Bottom Trawl survey using a species distribution model. There are estimated increases in the relative index of juvenile sablefish across various port groups associated with known strong sablefish recruitment: 2008, 2013, 2020 (no survey that year), and 2021. In recent years the relative abundance of juvenile sablefish has been highest around Newport.  In contrast, the highest relative abundance of juvenile sablefish were estimated to be around San Francisco and Bodega Bay in 2008 and Astoria and Tillamook in 2016.
```{r}
#| label: juvenile-by-port
#| echo: false
#| fig-cap: "Index of availability of juvenile sablefish (of 29 cm or less) to selected ports along the West Coast that fish heavily for Pacific hake. Availability was calculated within a 232 km radius. Annual values were scaled to 0-1 by dividing by the maximum observed biomass."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "GF-juv-sablefish-salmon-ports-1.png")
)

```

Stronger recruitment in the north, in recent years, has led to an increase in available biomass for fishers operating out of Astoria, but overall increases in sablefish biomass has led to general increases in available biomass across many ports.

```{r}
#| label: adult-by-port
#| echo: false
#| fig-cap: "Index of the availability of sablefish biomass to ports along the West Coast. Ports are Bellingham Bay (BLL), Astoria (AST), Charleston (Coos Bay, COS), Brookings (BRK), Crescent City (CRS), Eureka (ERK), Fort Bragg (BRG) and Morro Bay (MRO). Availablility was calculate based on a combination of locally available biomass and distance from port for the 75th quartile of fishing trips in 2015 using results from a species distribution model."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "sablefish-adults-avail-to-ports.png")
)

```


### Environmental Indices

The 2019 assessment of sablefish included an environmental index to inform the estimation of annual recruitment.  This assessment will also include an environmental index to inform the estimation of recent recruitment. The environmental index in this assessment will be based on a model with oceanographic predictors.The preliminary index is shown below. 

```{r}
#| label: glorys-index
#| echo: false
#| fig-cap: "Model fit to the data and model testing results. Black line is the model prediction ±1.0 standard error (dotted lines). While points are the recruitment deviations in log space from the stock assessment model. Yellow points are from jackknife analysis leaving out one year and refitting the model, and red points fit the best-fit model to 1993-2017 then predict 2018-2023."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "glorys-fit.png")
)

```


### Other Trawl Surveys

The assessment will include relative indices of abundance and composition data from historical U.S. West Coast trawl surveys: the Triennial Shelf Survey, AFSC Slope Survey, and NWFSC Slope Survey.  

Triennial Shelf Survey:

- The survey operated on a tri-annual basis starting in 1980 and ending in 2004.
- The survey sampled a depth range between 55-500 meters.
- The southern sampling range changed over time, extending to range between Point Conception in California to the U.S./Canadian border starting in 1992.
- The timing of sampling shifted to later in the summer months starting in 1995.

AFSC Slope Survey:

- The years of sampling used within West Coast groundfish assessments from this survey are 1997, 1999, 2000, and 2001.
- The survey sampled between Point Conception in California to the U.S./Canadian border.
- The survey sampled a depth range between 183-1,280 meters.

NWFSC Slope Survey:

- This survey operated annually between 1998 to 2002.
- The survey sampled between Point Conception in California to the U.S./Canadian border between 1998-2001 extended the southern sampling range into the area south of Point Conception in 2002.
- The survey sampled a depth range between 183-1,280 meters.

**Disclaimer: All data summaries and exploration presented here are preliminary and may not be indicative of the final data that will be incorporated in the 2025 assessment**



