```{r}
#| label: load-biology-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")

maturity_data <- readRDS(
  here::here("figures", "sablefish_maturity.rds")
) |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")
  )
maturity_at_age <- readRDS(
  here::here("figures", "ogives_for_assessment.rds")
)

wtatage_model <- r4ss::SS_readwtatage(
  file = file.path(data_proc, "wtatage_model_biomass_weighted.ss")) |>
  dplyr::filter(fleet == 1) |>
  dplyr::select(-seas, -bio_pattern, -birthseas) |>
  tidyr::pivot_longer(
    cols = 4:34,
    names_to = "age",
    values_to = "weight" 
  ) |>
  dplyr::filter(year < 2025, year > 2002) |>
  dplyr::mutate(
    sex = dplyr::case_when(sex == 1 ~ "Female", .default = "Male"),
    age = as.numeric(age),
    Year = as.factor(year)
  )

```

## General Approach to Biology

Similar to many groundfish species, sablefish, have variable growth and maturity along the U.S. West Coast, generally having latitudinal changes in biology with fish maturing early, growing faster, and reaching larger sizes in northern latitudes compared to southern latitudes. We have carefully considered how best to capture trends in biology in a coastwide model. When present, trends in biology are discussed below along with our proposed approach to estimate single coastwide parameters, when appropriate.

## Natural Mortality

Sablefish are a long-lived roundfish.  The maximum age observed varies by region. The reported maximum age for Alaska is 94 years, 113 years off Canada, and 102 years off the U.S. West Coast.  The maximum age observation off the West Coast was caught by the commercial fishery in 2006 and landed in Washington. The oldest aged sablefish by sex off the West Coast both arise from the fishery with an observation of 102 year old female and 98 year old male.

The 2011 and 2019 benchmark assessments for sablefish estimated sex-specific natural mortality within the model.  The 2019 assessment estimated a marginally higher natural mortality (0.076 yr$^{-yr}$) for females compared to males (0.067 yr$^{-yr}$), which approximately translates to a maximum age range between 71--80 years of age. This assessment will also attempt to estimate natural mortality within the assessment model. 

As sablefish age, both male and females move to deeper waters which likely reduces the availability of older fish to the fishery to some extent (i.e., although the oldest aged fish were sampled from the fishery). Age data off the West Coast start in 1983 with limited samples from the fishery and the Triennial Shelf Survey. However, these early data likely arise from the shelf area and may not have fully representative samples. Examining data from the surveys that sampled out to 1,280 meters (NWFSC WCGBT, NWFSC Slope, and AFSC Slope Surveys), the upper range of observed ages from these data sources by sex are:

- Females: 65 (22 observations of this age or higher) to a maximum observed age of 88 years; and
- Males: 70 (18 observations of this age or higher) to a maximum observed age of 91 years.

## Maturity

New data and analysis has been conducted to develop improved maturity estimates for this assessment. A total of `r dim(maturity_data)[1]` sablefish maturity samples were collected during the annual NWFSC West Coast Groundfish Bottom trawl survey in 2010, 2015, 2018, 2021, and 2022 (see table below). 

```{r}
#| label: maturity-samples
#| warning: false
#| echo: false
#| tbl-cap: "Maturity samples collected by the NWFSC West Coast Groundfish Bottom Trawl survey by year and area where north and south of 36 degress north latitude."
#| tbl-cap-location: margin

maturity_data |>
  dplyr::group_by(year, Area) |>
  dplyr::summarise(
    Samples = dplyr::n()
  ) |>
  tidyr::pivot_wider(
    names_from = Area,
    values_from = Samples,
    names_prefix = "Samples-"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Maturity samples by year and area."
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

These samples were evaluated for functional maturity that classifies maturity with developing oocytes as mature or immature. The maturity-at-age or -length estimates were evaluated north and south of 36$^\circ$ North latitude. Of the samples processed, sablefish were identified to be mature as early as age `r min(maturity_data[which(maturity_data$Area == "North" & maturity_data$functional_maturity == 1), "age"])` and `r min(maturity_data[which(maturity_data$Area == "South" & maturity_data$functional_maturity == 1), "age"])` in the north and south, respectively.  Generally, across samples, sablefish matured at younger ages more frequently in the north compared to the south.   

```{r}
#| label: maturity-data
#| echo: false
#| fig-cap: "Samples by age that were determined to be mature and immature north and south of 36 degrees north latitude. Immature fish have a functional maturity of 0 and mature fish have a functional maturity of 1. Blue vertical dashed line at age 7."
#| fig-cap-location: margin

ggplot(maturity_data, aes(x = age, y = functional_maturity)) +
  geom_point() +
  ylab("Functional Maturity") +
  geom_vline(xintercept = 7, color = "blue", linetype = 2) + 
  xlab("Age years") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  facet_wrap(Area ~ .)

```

Maturity-at-age was estimated using a spatiotemporal model that explored the impacts of temperature, temporal changes in maturity, day of year, and location. Only sample location was determined to be significant in the model with the maturity curve right shifted (e.g., less probability of being mature at age) compared to fish in the north. A coastwide maturity curve was estimated by weighting the area specific curves using a biomass weighted prediction grid from the NWFSC West Coast Groundfish Bottom Trawl survey. The assessment will be using the biomass weighted maturity curve (Spatial). 

```{r}
#| label: maturity-ogive
#| echo: false
#| fig-cap: "Estimated maturity-at-age by area and the biomass weighted coastwide (Spatial) maturity curve."
#| fig-cap-location: margin

ggplot(maturity_at_age, aes(x = age, y = p, color = Model)) +
  geom_line(linewidth = 1) +
  ylab("Probability Mature") +
  xlab("Age years") +
  scale_colour_viridis_d() + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"))

```

## Growth

```{r}
#| label: load-growth-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)

load(file.path(data, "data_survey_bio.rda"))
cols_to_keep <- c(
  "Project",
  "Year",
  "Sex",
  "Age_years", 
  "Length_cm", 
  "Weight_kg",
  "Depth_m", 
  "Latitude_dd"
)

survey_bio <- dplyr::bind_rows(
  data_survey_bio$nwfsc_combo |> dplyr::select(dplyr::all_of(cols_to_keep)), 
  data_survey_bio$nwfsc_slope |> dplyr::select(dplyr::all_of(cols_to_keep)),
  data_survey_bio$afsc_slope$age_data |> dplyr::select(dplyr::all_of(cols_to_keep)),
  data_survey_bio$triennial_early$age_data |> dplyr::select(dplyr::all_of(cols_to_keep)),
  data_survey_bio$triennial_late$age_data |> dplyr::select(dplyr::all_of(cols_to_keep))
) |>
  dplyr::mutate(
    state = dplyr::case_when(Latitude_dd > 46.25 ~ "WA", Latitude_dd < 42 ~ "CA", .default = "OR"),
    area = dplyr::case_when(Latitude_dd > 36 ~ "North", .default = "South")
  ) 

growth_cw <- nwfscSurvey::est_growth(
  dat = survey_bio |> 
    dplyr::filter(Sex != "U") |> dplyr::mutate(Age = Age_years),
  return_df = FALSE
)

growth_north <- nwfscSurvey::est_growth(
  dat = survey_bio |> 
    dplyr::filter(area == "North", Sex != "U") |> dplyr::mutate(Age = Age_years),
  return_df = FALSE
)
growth_south <- nwfscSurvey::est_growth(
  dat = survey_bio |> 
    dplyr::filter(area == "South", Sex != "U") |> dplyr::mutate(Age = Age_years),
  return_df = FALSE
)

growth <- dplyr::bind_rows(
  growth_cw$female_growth,
  growth_cw$male_growth,
  growth_north$female_growth,
  growth_north$male_growth,
  growth_south$female_growth,
  growth_south$male_growth
)
growth$sex <- rep(c("F", "M"), 3)
growth$area <- c(rep("Coastwide", 2), rep("North", 2), rep("South", 2))
```

Sablefish have dimorphic growth with females reaching larger sizes compared to males.  

```{r}
#| label: plot-growth-data
#| echo: false
#| warning: false
#| fig-cap: "Estimated length-at-age coastwide and by area for females (solid lines) and males (dashed lines) north and south of 36 degrees North latitude."
#| fig-cap-location: margin

ggplot(data = survey_bio, aes(x = Age_years, y = Length_cm, color = Sex)) +
  geom_point()  +
 xlim(0, 95) +
 ylim(0, 90) +
 theme_bw() +
 xlab("Age (yrs)") +
 ylab("Length (cm)") +
 scale_color_viridis_d() + 
 theme(axis.text = element_text(size = 13),
       axis.title = element_text(size = 13)) +
 facet_grid(area~.)

```

Length-at-age was estimated assuming von Bertenlaffy growth collected during various West Coast bottom trawl surveys (Triennial, NWFSC Slope, AFSC Slope, and NWFSC West Coast Groundfish Bottom Trawl survey). Additionally, fish north of 36$^\circ$ North latitude grow to larger sizes compared to fish in the south. Coastwide, sablefish reach an average maximum size around `r round(growth[which(growth$area == "Coastwide" & growth$sex == "F"), "Linf"], 0)` cm and `r round(growth[which(growth$area == "Coastwide" & growth$sex == "M"), "Linf"], 0)` cm for females and males, respectively. However, there is considerable variation in growth for sablefish where individual fish may grow to a larger or smaller maximum size.

```{r}
#| label: plot-growth
#| echo: false
#| warning: false
#| fig-cap: "Estimated length-at-age coastwide and by area for females (solid lines) and males (dashed lines) north and south of 36 degrees North latitude."
#| fig-cap-location: margin

data_to_plot <- survey_bio |>
  dplyr::rename_with(
    tolower
  ) |>
  dplyr::filter(
    sex != "U",
    !is.na(age_years),
    age_years >= 0,
    !is.na(length_cm),
    length_cm > 0
  )

xlims <- c(0, ceiling(max(data_to_plot[, "age_years"])))
ylims <- c(0, max(data_to_plot[, "length_cm"]))
growth$amin <- 0
growth$amax <- xlims[2]

lines_to_plot <- growth |>
  dplyr::group_by(area, sex) |>
  dplyr::reframe(
    k = K,
    Linf = Linf,
    L0 = L0,
    age_years = seq(amin, amax, 1),
    length_cm = Linf + (L0 - Linf) * exp(-k * age_years)
  )
label <- lines_to_plot |>
  dplyr::mutate(
    max_x = quantile(age_years, 0.60),
    multiplier = ifelse(sex == "Female", 0.2, 0.1)
  ) |>
  dplyr::group_by(area, sex) |>
  dplyr::summarize(
    label = paste0(
      "k = ", round(unique(k), 2), "; ",
      paste0("Lmin = ", round(unique(L0), 1)), "; ",
      paste0("Linf = ", round(unique(Linf), 1))
    ),
    x = unique(max_x),
    y = unique(max(length_cm)) * unique(multiplier)
  )

ggplot(lines_to_plot) + 
 geom_line(aes(x = age_years, y = length_cm, color = area, linetype = sex), linewidth = 1) +
 xlim(xlims[1], xlims[2]) +
 ylim(24, 70) +
 theme_bw() +
 xlab("Age (yrs)") +
 ylab("Length (cm)") +
 scale_color_viridis_d() + 
 theme(axis.text = element_text(size = 13),
       axis.title = element_text(size = 13))

```

```{r}
#| label: growth-continous
#| echo: false
#| fig-cap: "Latitudinal variation in sex-specific von Bertalanffy growth parameters (t0: length at age 0, k: growth coefficient, L∞: asymptotic length). Dashed horizontal line denotes 36.0° N. latitude. Predicted using the NWFSC West Coast Groundfish Bottom Trawl survey data between 2003 and 2024. The darker color represents females and the lighter color represents males."
#| fig-cap-location: margin

knitr::include_graphics(
  file.path("figures", "Fig4_Mod.png")
)

```

Previous assessments of sablefish have estimated time-invariant sex-specific growth within the assessment model. One of the recommendations from the 2019 Stock Assessment Review Panel was to explore using a time-varying annual weight-at-age approach in future assessments.  This approach utilizes tools to estimate the average weight of fish by sex and age, that is then fixed in the model. The advantage of this approach is that it can capture annual variations in growth without adding additional parameters within the model. This approach is used in the U.S./Canadian assessment of Pacific hake. This alternative approach for growth is discussed below.

## Weight-at-Age

Previous assessments of sablefish estimated a constant sex-specific growth curve (length-at-age) within the assessment model informed by length and age data. The 2019 STAR panel noted that the raw data for sablefish seemed to show considerable growth variability across time that was inconsistent with constant growth by sex. The STAR panel recommend future assessments explore replacing the assumption of constant growth with a weight-at-age approach. Traditionally the assessment model converts the growth curve relationship between length-at-age and the weight-at-length into a weight-at-age for all model calculations. Inputting an externally estimated weight-at-age relationship circumvents the need for a growth curve within the assessment model. Additionally, this approach allows for the definition of annual weight-at-age which can account for changes in growth based upon oceanographic, competition of resources, etc.

We have explored estimating weight-at-age by year, cohort, sex, and area where area is whether the sample was collected north or south of 36$^\circ$ N. latitude. Estimates were biomass weighted by area and combined for a coastwide weight-at-age by sex and year.     

```{r}
#| label: plot-tv-wtatage
#| echo: false
#| warning: false
#| fig-cap: "Biomass weighted estimated weight-at-age by year and sex."
#| fig-cap-location: margin

ggplot() +
  geom_line(data = wtatage_model, aes(x = age, y = weight, color = Year), linewidth = 1.0) +
  scale_color_viridis_d() +
  xlab("Age (years)") + ylab("Weight (kg)") +
  theme_bw() +
  theme(
    text = element_text(size = 15),
    axis.text = element_text(size = 15)
  ) +
  facet_grid(sex~.)

```