```{r}
#| label: load-biology-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")
data_raw <- file.path(project_dir , "data-raw")
maturity_data <- readRDS(
  file.path(data_raw, "maturity", "sablefish_maturity.rds")
) |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")
  )
maturity_at_age <- readRDS(
  file.path(data_raw, "maturity", "ogives_for_assessment.rds")
)
```

## Natural Mortality

## Maturity

New data and analysis has been conducted to develop improved maturity estimates for this assessment. A total of `r dim(maturity_data)[1]` sablefish maturity samples were collected during the annual NWFSC West Coast Groundfish Bottom trawl survey in 2010, 2015, 2018, 2021, and 2022 (see table below). 

```{r}
#| label: maturity-samples
#| warning: false
#| echo: false
#| tbl-cap: "Maturity samples collected by the NWFSC West Coast Groundfish Bottom Trawl survey by year and area where north and south of 36 degress north latitude."
#| tbl-cap-location: margin

maturity_data |>
  dplyr::group_by(year, Area) |>
  dplyr::summarise(
    Samples = dplyr::n()
  ) |>
  tidyr::pivot_wider(
    names_from = Area,
    values_from = Samples,
    names_prefix = "Samples-"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Maturity samples by year and area."
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20,
    table.align = "right"
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

These samples were evaluated for functional maturity that classifies maturity with developing oocytes as mature or immature. The maturity-at-age or -length estimates were evaluated north and south of 36$\circ$ North latitude.     

```{r}
#| label: maturity-data
#| echo: false
#| fig-cap: "Samples by age that were determined to be mature and immature north and south of 36 degrees north latitude. Immature fish have a functional maturity of 0 and mature fish have a functional maturity of 1. Blue vertical dashed line at age 7."
#| fig-cap-location: margin

ggplot(maturity_data, aes(x = age, y = functional_maturity)) +
  geom_point() +
  ylab("Functional Maturity") +
  geom_vline(xintercept = 7, color = "blue", linetype = 2) + 
  xlab("Age years") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  facet_wrap(Area ~ .)

```




```{r}
#| label: maturity-ogive
#| echo: false
#| fig-cap: "Add caption"
#| fig-cap-location: margin

ggplot(maturity_at_age, aes(x = age, y = p, color = Model)) +
  geom_line(linewidth = 1) +
  ylab("Probability Mature") +
  xlab("Age years") +
  scale_colour_viridis_d() + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"))

```



## Growth

## Weight-at-Age

## Stock Structure


