```{r}
#| label: load-biology-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")
data_raw <- file.path(project_dir , "data-raw")
maturity_data <- readRDS(
  file.path(data_raw, "maturity", "sablefish_maturity.rds")
) |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")
  )
maturity_at_age <- readRDS(
  file.path(data_raw, "maturity", "ogives_for_assessment.rds")
)
```

## General Approach to Biology

Similar to many groundfish species, sablefish, have variable growth and maturity along the U.S. West Coast, generally having latitudinal changes in biology with fish maturing early, growing faster, and reaching larger sizes in northern latitudes compared to southern latitudes. We have carefully considered how best to capture trends in biology in a coastwide model. When present, trends in biology are discussed below along with our proposed approach to estimate single coastwide parameters, when appropriate.

## Natural Mortality

Sablefish are a long-lived roundfish.  The maximum age observed varies by region. The reported maximum age for Alaska is 94 years, 113 years off Canada, and 102 years off the U.S. West Coast caught by the commercial fishery in 2006 landed in Washington. The 2011 and 2019 benchmark assessments for sablefish estimated sex-specific natural mortality with female being estimated have marginally higher natural mortality (0.076 yr$^{-yr}$) compared to males (0.062 yr$^{-yr}$) that approximately translate to a maximum age range between 70--87 years of age. This assessment will also attempt to estimate natural mortality within the assessment model. 

## Maturity

New data and analysis has been conducted to develop improved maturity estimates for this assessment. A total of `r dim(maturity_data)[1]` sablefish maturity samples were collected during the annual NWFSC West Coast Groundfish Bottom trawl survey in 2010, 2015, 2018, 2021, and 2022 (see table below). 

```{r}
#| label: maturity-samples
#| warning: false
#| echo: false
#| tbl-cap: "Maturity samples collected by the NWFSC West Coast Groundfish Bottom Trawl survey by year and area where north and south of 36 degress north latitude."
#| tbl-cap-location: margin

maturity_data |>
  dplyr::group_by(year, Area) |>
  dplyr::summarise(
    Samples = dplyr::n()
  ) |>
  tidyr::pivot_wider(
    names_from = Area,
    values_from = Samples,
    names_prefix = "Samples-"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Maturity samples by year and area."
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

These samples were evaluated for functional maturity that classifies maturity with developing oocytes as mature or immature. The maturity-at-age or -length estimates were evaluated north and south of 36$^\circ$ North latitude. Of the samples processed, sablefish were identified to be mature as early as age `r min(maturity_data[which(maturity_data$Area == "North" & maturity_data$functional_maturity == 1), "age"])` and `r min(maturity_data[which(maturity_data$Area == "South" & maturity_data$functional_maturity == 1), "age"])` in the north and south, respectively.  Generally, across samples, sablefish matured at younger ages more frequently in the north compared to the south.   

```{r}
#| label: maturity-data
#| echo: false
#| fig-cap: "Samples by age that were determined to be mature and immature north and south of 36 degrees north latitude. Immature fish have a functional maturity of 0 and mature fish have a functional maturity of 1. Blue vertical dashed line at age 7."
#| fig-cap-location: margin

ggplot(maturity_data, aes(x = age, y = functional_maturity)) +
  geom_point() +
  ylab("Functional Maturity") +
  geom_vline(xintercept = 7, color = "blue", linetype = 2) + 
  xlab("Age years") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  facet_wrap(Area ~ .)

```

Maturity-at-age was estimated using a spatiotemporal model that explored the impacts of temperature, temporal changes in maturity, day of year, and location. Only sample location was determined to be significant in the model with the maturity curve right shifted (e.g., less probability of being mature at age) compared to fish in the north. A coastwide maturity curve was estimated by weighting the area specific curves using a biomass weighted prediction grid from the NWFSC West Coast Groundfish Bottom Trawl survey. The assessment will be using the biomass weighted maturity curve (Spatial). 

```{r}
#| label: maturity-ogive
#| echo: false
#| fig-cap: "Estimated maturity-at-age by area and the biomass weighted coastwide (Spatial) maturity curve."
#| fig-cap-location: margin

ggplot(maturity_at_age, aes(x = age, y = p, color = Model)) +
  geom_line(linewidth = 1) +
  ylab("Probability Mature") +
  xlab("Age years") +
  scale_colour_viridis_d() + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"))

```

## Growth

## Weight-at-Age

## Stock Structure


