```{r}
#| label: load-biology-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")
data_raw <- file.path(project_dir , "data-raw")
maturity_data <- readRDS(
  file.path(data_raw, "maturity", "sablefish_maturity.rds")
) |>
  dplyr::mutate(
    Area = dplyr::case_when(latitude >= 36 ~ "North", .default = "South")
  )
maturity_at_age <- readRDS(
  file.path(data_raw, "maturity", "ogives_for_assessment.rds")
)
```

## Natural Mortality

## Maturity

New data and analysis has been conducted to develop improved maturity estimates for this assessment. A total of `r dim(maturity_data)[1]` sablefish maturity samples were collected during the annual NWFSC West Coast Groundfish Bottom trawl survey in 2010, 2015, 2018, 2021, and 2022 (see table below). 

```{r}
#| label: maturity-samples
#| warning: false
#| echo: false
#| tbl-cap: "Maturity samples collected by the NWFSC West Coast Groundfish Bottom Trawl survey by year and area where north and south of 36 degress north latitude."
#| tbl-cap-location: margin

maturity_data |>
  dplyr::group_by(year, Area) |>
  dplyr::summarise(
    Samples = dplyr::n()
  ) |>
  tidyr::pivot_wider(
    names_from = Area,
    values_from = Samples,
    names_prefix = "Samples-"
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Maturity samples by year and area."
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

These samples were evaluated for functional maturity that classifies maturity with developing oocytes as mature or immature. The maturity-at-age or -length estimates were evaluated north and south of 36$^\circ$ North latitude. Of the samples processed, sablefish were identified to be mature as early as age `r min(maturity_data[which(maturity_data$Area == "North" & maturity_data$functional_maturity == 1), "age"])` and `r min(maturity_data[which(maturity_data$Area == "South" & maturity_data$functional_maturity == 1), "age"])` in the north and south, respectively.  Generally, across samples, sablefish matured at younger ages more frequently in the north compared to the south.   

```{r}
#| label: maturity-data
#| echo: false
#| fig-cap: "Samples by age that were determined to be mature and immature north and south of 36 degrees north latitude. Immature fish have a functional maturity of 0 and mature fish have a functional maturity of 1. Blue vertical dashed line at age 7."
#| fig-cap-location: margin

ggplot(maturity_data, aes(x = age, y = functional_maturity)) +
  geom_point() +
  ylab("Functional Maturity") +
  geom_vline(xintercept = 7, color = "blue", linetype = 2) + 
  xlab("Age years") +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  facet_wrap(Area ~ .)

```

Maturity-at-age was estimated using a spatiotemporal model that explored the impacts of temperature, temporal changes in maturity, day of year, and location. Only sample location was determined to be significant in the model with the maturity curve right shifted (e.g., less probability of being mature at age) compared to fish in the north. A coastwide maturity curve was estimated by weighting the area specific curves using a biomass weighted prediction grid from the NWFSC West Coast Groundfish Bottom Trawl survey. 

```{r}
#| label: maturity-ogive
#| echo: false
#| fig-cap: "Estimated maturity-at-age by area and the biomass weighted coastwide (Spatial) maturity curve."
#| fig-cap-location: margin

ggplot(maturity_at_age, aes(x = age, y = p, color = Model)) +
  geom_line(linewidth = 1) +
  ylab("Probability Mature") +
  xlab("Age years") +
  scale_colour_viridis_d() + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"))

```



## Growth

## Weight-at-Age

## Stock Structure


