```{r}
#| label: load-fishery-data
#| echo: false
#| warning: false
library(ggplot2)
library(dplyr)
library(gt)
# this needs to be fixed to use relative directories
# project_dir <- here::here()
project_dir <- "C:/Assessments/2025/sablefish_2025"
data_proc <- file.path(project_dir, "data-processed")
data <- file.path(project_dir , "data")

load(file.path(data, "data_commercial_landings.rda"))
load(file.path(data, "data_commercial_discards.rda"))
landings <- dplyr::bind_rows(
  data_commercial_landings,
  data_commercial_discards) |>
  dplyr::mutate(
    Fleet = dplyr::case_when(
      fleet == 1 ~ "Trawl",
      fleet == 2 ~ "Hook & Line",
      fleet == 3 ~ "Pot",
      fleet == 4 ~ "Trawl Discard",
      fleet == 5 ~ "Hook & Line Discard",
      fleet == 6 ~ "Pot Discard",)
  ) 

load(file.path(data, "data_commercial_bds.rda"))
bds_data <- data_commercial_bds |> 
  as_tibble() |>
  mutate(
    area = case_when(
      state %in% c("WA", "OR") ~ state,
      TRUE ~ paste(area, state)
    ),
    period = case_when(
      year > 2018 ~ "2019-2024",
      year < 2011 ~ "1980-2010",
      TRUE ~ "2011-2018"
    ),
    area = factor(area, levels=c("South CA", "North CA", "OR", "WA"))
  )

```

## General Updates 

Updates are provided below on the available data and modeling decisions that may have changed since the pre-assessment data workshop.

### Fishery Data

#### Fleets Proposed for the Assessment

Since the pre-assessment data workshop, the stock assessment team has modified the landings and the proposed model structure to address incorrect assumptions from previous assessments and to support a more stable model. Previous assessments of West Coast sablefish have added historical foreign fleet removals and domestic removals together into gear-specific landings fleets. Previous assessments then estimated additional mortality from discarding (i.e., using estimates of retention and the gear-specific discard mortality rates) for all model years based on the combined catch streams from the foreign and domestic fisheries.  When reviewing the historical documentation that provided the estimates of historical removals by foreign fleets in U.S. waters, it was determined that the reported totals by gear type and year represented total catches (e.g., landings plus discards) meaning it would be inaccurate to estimate additional discard mortality from these catch totals which was done in previous assessments.  

The other factor that had led to the stock assessment team modifying the catch streams has been driven by model stability and transparency.  Groundfish assessments have applied either of the following approaches to account for total mortality (landings plus dead discards): 1) estimate selectivity and retention with assumptions about discard mortality within the model using available discard data (discard rates and lengths or ages of discarded fish) or 2) externally calculate discard mortality based on landings and data about discarding practices where the calculated discard mortality is added to the landings or include in separate discard fleets in the model. These two approaches, when the models are parameterized in a comparable fashion, should result in very similar model estimates. Previous assessments of sablefish have opted to estimate selectivity and retention within the model. However, during model development, the stock assessment team determined that this model structure tended to be more unstable, creating challenges in estimating discard mortality that aligned with estimates by gear from the Groundfish Expanded Multiyear Mortality (GEMM) report, and that this model was less transparent to evaluate when it was producing reasonable estimates of selectivity and retention. Given these challenges, the stock assessment team, opted to move to a model structure with separate discard fleets.  

The new fleets used in the model will increase from three (trawl, hook & line, and pot) to a total of six fishery fleets with the additional three fleets that represent estimated discard mortality by gear group.  The discard mortality was estimated based upon the gear using historical discard studies (e.g., Pikitch) and current discard information collected by the West Coast Groundfish Observer Program (WCGOP). In summary, the proposed fleets are now:

- Trawl: Includes domestic historical and current landings, historical foreign fleet total mortality, and total mortality estimates from the at-sea fishery.
- Hook & Line: Includes domestic historical and current landings, historical foreign fleet total mortality, and recreational total mortality.
- Pot: Includes domestic historical and current landings and historical foreign fleet total mortality.
- Trawl Discards: Externally estimated discard mortality based on the domestic historical and current landings informed by discard rate data (Pikitch and WCGOP) with an assumed mortality rate of 50%.
- Hook & Line Discards: Externally estimated discard mortality based on the domestic historical and current landings informed by discard rate data (WCGOP) with an assumed mortality rate of 20%.
- Pot Discards: Externally estimated discard mortality based on the domestic historical and current landings informed by discard rate data (WCGOP) with an assumed mortality rate of 20%.

The revised removals that will be used in the assessment are provided below.

```{r}
#| label: landings-all-years
#| echo: false
#| warning: false
#| fig-cap: "Coastwide landings and discads (mt) of sablefish by gear."
#| fig-cap-location: margin

ggplot(landings, aes(x = year, y = catch_mt, fill = Fleet)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("Year") + ylab("Landings & Discards (mt)") +
  scale_y_continuous(
    labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  scale_fill_viridis_d()

```


```{r}
#| label: landings-recent
#| echo: false
#| warning: false
#| fig-cap: "Coastwide landings and discards (mt) of sablefish by gear since 1981. "
#| fig-cap-location: margin

ggplot(landings |> dplyr::filter(year >= 1981), aes(x = year, y = catch_mt, fill = Fleet)) +
  geom_bar(stat = 'identity') +
  theme_bw() +
  xlab("Year") + ylab("Landings & Discards (mt)") +
  scale_y_continuous(
    labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  scale_fill_viridis_d()

```

```{r}
#| label: landings-table
#| warning: false
#| echo: false
#| tbl-cap: "Coastwide landings (mt) by gear and total mortality for the last ten years.  Note, the landings for 2024 are incomplete."
#| tbl-cap-location: margin

landings |>
  dplyr::filter(year >= 2015) |>
  dplyr::group_by(year, Fleet) |>
  dplyr::summarise(
    mt = sum(catch_mt),
  ) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    total = sum(mt)
  ) |>
  dplyr::select(year, Fleet, mt, total) |>
  tidyr::pivot_wider(
    names_from = Fleet,
    values_from = mt
  ) |>
  dplyr::relocate(total, .after = `Trawl Discard`) |>
  dplyr::rename(
    Year = year,
    `Total Mortality (mt)` = total
  ) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Coastwide landings and discards (mt) by gear and total landings for the last ten years. "
  ) |>
  fmt_number(
    columns = 2:8,
    decimals = 0
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20,
    table.align = "right"
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 

```

#### Fishery Age Data

At the time of the pre-assessment data workshop, recent ages were not available for discussion.  The Cooperative Ageing Lab has since completed ageing work to support this assessment.  A table with the number of ages by fleet for the most recent 10 years and age distribution plots representing all available age data are provided below.

```{r}
#| label: nsamples-age
#| echo: false
#| warning: false
#| tbl-cap: "Number of available ages over time, grouped by gear type (TWL = trawl, HKL = hook-and-line, POT = pot) and state. Note, this does not include unread otoliths."
#| tbl-cap-location: margin

bds_data |> 
  dplyr::filter(year >= 2015) |>
  dplyr::group_by(year, geargroup) |>
  dplyr::summarise(Ages = sum(!is.na(age_years))) |>
  dplyr::mutate(
    geargroup = dplyr::case_when(
      geargroup == "TWL" ~ "Trawl", geargroup == "HKL" ~ "Hook & Line", .default = "Pot")
  ) |>
  tidyr::pivot_wider(
    names_from = geargroup,
    values_from = Ages
  ) |>
  dplyr::rename(Year = year) |>
  dplyr::ungroup() |>
  gt() |>
  tab_header(
    title = "Number of age samples by gear type for the last 10 years. "
  ) |>
  cols_align(
    align = "center"
  ) |>
  tab_options(
    table.font.size = 20,
    table.align = "right"
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "white"),
      cell_text(color = "black")
    ),
    locations = cells_body(
      columns = everything()
    )
  ) 
```

```{r}
#| label: age-by-gear-region
#| echo: false
#| warning: false
#| fig-cap: "Distribution of all age samples between 1986-2024 by gear type and state."
#| fig-cap-location: margin

bds_data |>
  group_by(area, geargroup) |>
  mutate(area = factor(area)) |>
  select(sex, age_years, area, geargroup) |>
  ggplot(aes(x = age_years, fill = area))+
    ggridges::geom_density_ridges(aes(y = geargroup), alpha = 0.5, scale = 1)+
    labs(y = "Gear", x = "Age (years)", fill = "Area")+
    scale_x_continuous(limits = c(0, 30), expand = c(0, 0))+
    scale_fill_viridis_d()+
    theme_bw()
```


### Survey Data

#### AFSC Slope Survey

Since the pre-assessment data workshop, the stock assessment team investigated the data from the AFSC Slope survey in depth.  During these investigations it was determined that aged fish for sablefish did not appear to be a representative sample of the lengths observed (e.g., fish identified for otolith collection were not selected randomly from all fish caught). This same issue was also identified for this survey in the most recent assessment of Dover sole. 

Additionally, this survey had an unusually high rate of positive tows for sablefish (e.g., ~ 98% of tows observed sablefish), particularly compared to the NWFSC Slope survey (~ 86%) that was conducted across similar years and depths. Given the concerns about the data collected by the AFSC Slope survey for sablefish, and the fact that there was a comparable survey data set during these years, the stock assessment team has decided to not include data from the AFSC Slope survey in this assessment.

#### Triennial Shelf Survey

The age data collected by the Triennial Shelf Survey was also evaluated to determine if the ages appeared to be a random sample of the lengths observed and whether the age data were collected across the survey range.  The number of tows with available age data for sablefish was compared to the tows with measured lengths. There were a select number of years where the age data were collected from very few tows, did not appear to be a random sample of the lengths measured, and/or were collected across very limited areas of the coast. The stock assessment team opted to use age data from 1983, 1989, 1992, 1995, and 2004 and to not include the age data collected in 1986, 1998, and 2001. 


